\documentclass[../Main.tex]{subfiles}

\begin{document}
\section{Velocity Potential}
In Chapter 2 we saw that if at $t = 0$, $\vec{\omega} = \vec0$, then $\vec{\omega} = \vec0$ at all time. That is, irrotational flow remains irrotational.

If $\vec{\omega} = \nabla \times \vec{u} = \vec0$, then we know from the course IA Vector Calculus that:
\begin{equation*}
    \vec{u}(\vec{x}, t) = \nabla \phi(\vec{x}, t)
\end{equation*}
where $\phi$ is the \underline{velocity potential}.
\begin{remark}
    We use a positive grad here, unlike, for example, $\bdforce = - \nabla \chi$.
\end{remark}
By applying the incompressibility of the flow, we get $\nabla \cdot (\nabla \phi) = 0$. We also apply kinematic boundary conditions to get the problem:
\begin{equation}
    \begin{cases}
        \nabla^2 \phi = 0 & \vec{x} \in \Omega \\
        \frac{\partial \phi}{\partial \vec{n}} = \vec{U} \cdot \vec{n} & \vec{x} \in \partial \Omega
    \end{cases}
    \label{eqnPotentialLaplace}
\end{equation}
Therefore, the Euler equation reduces to the Laplace equation (with Neumann boundary conditions). We use the Euler equation after we have found the flow in order to get the pressure.
\section{Simple Examples}
For simple geometries, we can build solutions using separable solutions to the Laplace Equation in suitable coordinate systems.

The simplest example we could think of is uniform flow $\vec{u} = \vec{U}$. Then $\phi = \vec{U} \cdot \vec{x}$.

\begin{example}[Spherical coordinates]
    The most general axisymmetric solution to Laplace's equation is:
    \begin{equation*}
        \phi = \sum_{n=0}^{\infty} \left(A_n r^n + B_n r^{-1-n}\right) P_n(\cos(\theta))
    \end{equation*}
    where $P_n$ are the Legendre polynomials.

    Usually we only need a small number of modes.
    \begin{itemize}
        \item The $B$ term when $n = 0$ is $\phi = B / r$. This gives a \underline{3D point source}, with velocity $\vec{u} = -\frac{B}{r^2} \vec{e_r}$. This is a radial flow with constant volume flux through a sphere:
            \begin{equation*}
                \int_{B_R(\vec0)} \vec{u} \cdot \vec{n} dS = -\frac{B}{R^2} \times 4\pi R^2 = -4\pi B
            \end{equation*}
            we define this to be the volume flux $q$, so:
            \begin{equation*}
                \phi = -\frac{q}{4\pi r}
            \end{equation*}
        \item The $B$ term when $n = 1$ is:
            \begin{equation*}
                \phi = B \frac{\cos(\theta)}{r^2}
            \end{equation*}
            this is \underline{dipole flow}, where a source and a sink are very close to each other.
    \end{itemize}
\end{example}
\begin{example}[Cylindrical coordinates]
    Consider cylindrical coordinates with no $z$ dependence (equivalent to 2D polar coordinates).

    The most general solution to Laplace's Equation is:
    \begin{equation*}
        \rho = C + A_0 \log(r) + B_0 \theta + \sum_{n=1}^{\infty} (A_n r^n + B_n r^{-n}) (C_n cos(n\theta) + D_n \sin(n\theta))
    \end{equation*}
    Since this is only defined up to a constant, we disregard $C$.
    
    We can take just a few nodes to give:
    \item With only the $A_0$ term,
        \begin{equation*}
            \phi = \frac{q}{2\pi}\log(r) \implies \vec{u} = \frac{q}{2\pi r} \vec{e_r}
        \end{equation*}
        which is a \underline{2D point source}.
    \item With only the $B_0$ term,
        \begin{equation*}
            \phi = \frac{\kappa}{2\pi}\theta \implies \vec{u} = \frac{\kappa}{2\pi r} \vec{e_\theta}
        \end{equation*}
        then this is circular motion around the origin. The vorticity is zero except at $0$ (vorticity is a delta function). Therefore this flow is invalid if we include the origin. 
    \item $\phi = Ur\cos(\theta)$ is a solution. %TODO: How?
    \item The \underline{2D dipole} is given by:
        \begin{equation*}
            \phi = A \frac{\cos(\theta)}{r}
        \end{equation*}
\end{example}
\section{Uniform Flow Past a Sphere or Cylinder}
\subsection{Flow Past a Sphere}
\begin{figure}
    \centering
    \begin{tikzpicture}[scale=1]
        \draw (0, 0) circle[radius=2] -- (30:2) node[pos=0.5, anchor=south] {$a$};

        \draw[->, blue] (-4.5, -0.3) -- (-4, -0.3) node[pos=0.5, above] {$U$};

        \draw[blue, domain=-5:5] plot (\x, {1.5 + (1 / 1.3) * exp(-\x * \x / 2.6)});
        \draw[blue, domain=-5:5] plot (\x, {2 + (1 / 1.8) * exp(-\x * \x / 3.6)});
        \draw[blue, domain=-5:5] plot (\x, {-1.5 - (1 / 1.3) * exp(-\x * \x / 2.6)});
        \draw[blue, domain=-5:5] plot (\x, {-2 - (1 / 1.8) * exp(-\x * \x / 3.6)});

        \begin{scope}[shift={(3,0)}]
            \draw[->] (0, 0) -- (30:1) node[pos=0.5, anchor=south] {$r$} node[pos=0.8, anchor=south west] {$\vec{x}$};
            \draw[->] (0, 0) -- (1, 0) node[right] {$\vec{e_z}$};
            \draw (0.5, 0) arc[radius=0.5, start angle=0, end angle=30] node[pos=0.7, right] {$\theta$};
        \end{scope}
    \end{tikzpicture}
    \caption{Diagram of flow past a sphere}
    \label{figFlowSphere}
\end{figure}
Consider figure~\ref{figFlowSphere}. We want to solve $\nabla^2 \phi = 0$ with the boundary conditions:
\begin{align*}
    \vec{u} &\to U \vec{e_z} \text{ far from the sphere} \\
    \vec{u} \cdot \vec{n} &= \vec{U} \cdot \vec{n} \text{ (KBC)}
\end{align*}
in terms of the velocity potential $\phi$,
\begin{align*}
    \phi(r, \theta) &\to U r \cos(\theta) \text{ as $r\to\infty$} \\
    \frac{\partial \phi}{\partial r}(a, \theta) &= 0 
\end{align*}
Then we try a solution of the form $\phi = f(r) P_1(\cos(\theta)) = f(r) \cos(\theta)$. We recall the general solution to the Laplace equation and suppose that $f(r)$ is of the form:
\begin{equation*}
    f(r) = Ar + B r^{-2}
\end{equation*}
We know that this solves the Laplace equation, so it only remains to check the boundary conditions. Flow at infinity gives $U = A$, and the kinematic boundary condition gives:
\begin{align*}
    \frac{\partial \phi}{\partial r} &= \left[U - \frac{2B}{r^3}\right]\cos(\theta) \\
    \left.\frac{\partial \phi}{\partial r}\right|_{r = a} &= \left[U - \frac{2B}{a^3}\right]\cos(\theta) = 0 \\
    \therefore 2B &= Ua^3 \\
    \therefore B &= \frac{Ua^3}{2}
\end{align*}
And so our solution for $\phi$ is:
\begin{equation*}
    \phi = U \cos(\theta) \left[r + \frac{a^3}{2r^2}\right]
\end{equation*}
which gives a solution for $\vec{u}$ by taking a gradient:
\begin{equation*}
    \vec{u} = U \cos(\theta) \left(1 - \frac{a^3}{r^3}\right)\vec{e_r} - U \sin(\theta) \left(1 + \frac{a^3}{2r^3}\right)\vec{e_\theta}
\end{equation*}
\subsection{Uniform Flow Past a Cylinder with Circulation}
\begin{figure}
    \centering
    \begin{tikzpicture}[>=stealth, scale=0.7]
    
        % Left Diagram
        % Object
        \draw[very thick] (0,0) circle (1);
        
        % Streamlines
        \begin{scope}[blue!80!black, thick]
        \draw (-4.5, 1.0) .. controls (-2.5, 1.0) and (-1.5, 1.5) .. (0, 1.5) .. controls (1.5, 1.5) and (2.5, 1.0) .. (4.5, 1.0);
        \draw (-4.5, 0.3) .. controls (-2.5, 0.3) and (-1.2, 1.05) .. (0, 1.05) .. controls (1.2, 1.05) and (2.5, 0.3) .. (4.5, 0.3);
        \draw (-4.5, -0.3) .. controls (-2.5, -0.3) and (-1.2, -1.05) .. (0, -1.05) .. controls (1.2, -1.05) and (2.5, -0.3) .. (4.5, -0.3);
        \draw (-4.5, -1.0) .. controls (-2.5, -1.0) and (-1.5, -1.5) .. (0, -1.5) .. controls (1.5, -1.5) and (2.5, -1.0) .. (4.5, -1.0);
        \end{scope}
        
        % Annotations
        \node[font=\Large] at (1.5, 2.2) {$|\vec{u}| = 2U$};
        \draw[->, thick] (1.0, 1.9) .. controls (0.8, 1.5) and (0.4, 1.2) .. (0.05, 1.05);
        
        \node[font=\Large] at (1.5, -2.2) {$|\vec{u}| = 2U$};
        \draw[->, thick] (1.0, -1.9) .. controls (0.8, -1.5) and (0.4, -1.2) .. (0.05, -1.05);
        
        
        % Right Diagram
        \begin{scope}[xshift=10cm]
        % Object
        \draw[very thick] (0,0) ellipse[radius=1];
        
        % Streamlines
        \begin{scope}[blue!80!black, thick]
        \draw (-4.5, 1.0) .. controls (-2.5, 1.2) and (-1, 1.9) .. (0, 1.9) .. controls (1, 1.9) and (2.5, 1.0) .. (4.5, 1.0);
        \draw (-4.5, 0.3) .. controls (-2.5, 0.5) and (-1, 1.2) .. (0, 1.2) .. controls (1, 1.2) and (2.5, 0.3) .. (4.5, 0.3);
        \draw (-4.5, -0.4) .. controls (-2.5, -0.4) and (-1, -1.2) .. (0, -1.2) .. controls (1, -1.2) and (2.5, -0.4) .. (4.5, -0.4);
        \end{scope}
        
        % Annotations
        \node[font=\Large] at (-0.5, 2.5) {$|\vec{u}| = U$};
        \draw[->, thick] (0.5, 2.5) .. controls (2.0, 2.5) and (1.5, 1.0) .. (0.3, 0.95);
        
        \node[font=\Large] at (-2.0, -2.5) {$|\vec{u}| = 3U$};
        \draw[->, thick] (-1.2, -2.1) .. controls (-0.8, -1.8) and (-0.6, -1.4) .. (-0.5, -0.9);
        \end{scope}
    
    \end{tikzpicture}
    \caption{Diagram of flow past a cylinder}
    \label{figFlowCylinderComparison}
\end{figure}
We consider cylindrical polar coordinates and solve $\nabla^2 \phi = 0$. The boundary conditions are:
\begin{align*}
    \phi &\xrightarrow[r \to \infty]{} Ur\cos(\theta) \\
    \frac{\partial \phi}{\partial r}(a, \theta) = 0 \\
    \oint_{B_r(\vec0)} \vec{u} \cdot \vec{dx} &= \kappa
\end{align*}
We cannot remove the circulation flow in this case, because we let the cylinder have infinite length along the $z$ axis.

The solution is:
\begin{align*}
    \phi &= \frac{\kappa}{2\pi}\theta + U\cos(\theta) \left[r + \frac{a^2}{r}\right] \\
    \therefore \vec{u} &= U\cos\left(\theta\right) (1 - \frac{a^2}{r^2}) \vec{e_r} - U\sin(\theta) \left(1 + \frac{a^2}{r^2}\right)\vec{e_\theta} + \frac{\kappa}{2\pi r} \vec{e_\theta}
\end{align*}
Importantly, we only have symmetry along the $x$ axis when $\kappa = 0$. See figure~\ref{figFlowCylinderComparison} for an illustration of this. An important consequence is that there will be a difference in pressure on either side of the cylinder, and so the fluid exerts a force on the cylinder.
\section{Pressure in a Potential Flow}
The Euler equation for pressure is:
\begin{equation*}
    \rho \frac{D \vec{u}}{D t} = - \nabla p + \bdforce
\end{equation*}
which is possibly nonlinear in $p$, even when we have found $\vec{u}$.

Now assume: $\bdforce = - \nabla \chi$, so the body forces are conservative, and $\rho$ is constant.
\begin{align*}
    0 &= \nabla p + \nabla \chi + \rho \left(\frac{\partial \vec{u}}{\partial t} + (\vec{u} \cdot \nabla)\vec{u}\right) \\
    0 &= \nabla p + \nabla \chi + \rho \frac{\partial (\nabla \phi)}{\partial t} + \rho \nabla \left(\frac12 |\vec{u}|^2\right) - \vec{u} \times \underbrace{\vec{\omega}}_{0} \\
    0 &= \nabla p + \nabla \chi + \nabla\left(\rho\frac{\partial \phi}{\partial t}\right) + \nabla \left(\rho\frac12 |\vec{u}|^2\right) \\
    0 &= \nabla\left( p + \chi + \rho\frac{\partial \phi}{\partial t} + \frac12\rho |\vec{u}|^2\right)
\end{align*}
and this is equivalent to:
\begin{equation}
    \rho \frac{\partial \phi}{\partial t} + \frac12 \rho |\nabla \phi|^2 + p + \chi = f(t)
    \label{eqnBernoulliUnsteady}
\end{equation}
where $f$ is any function of time. This is the \underline{Bernoulli Equation for}\\\underline{Unsteady Potential Flow} (unsteady Bernoulli).

This has key differences from \eqnref{eqnBernoulliSteady}:
\begin{enumerate}
    \item \eqnref{eqnBernoulliSteady} requires steady flow, but permits vorticity, \eqnref{eqnBernoulliUnsteady} requires irrotational flow, but permits time dependence;
    \item both require conservative forces
    \item \eqnref{eqnBernoulliSteady} is just a consequence of the Euler momentum equation~\ref{eqnEulerMomDiv}, because we took a scalar product; \eqnref{eqnBernoulliUnsteady} is equivalent to the Euler equation under the given assumptions.
\end{enumerate}
\section{Forces on a Translating Sphere or Cylinder}
\subsection{Steady Flow on a Translating Sphere}\label{secTranslatingSphere}
Consider steady flow, and move in the reference frame of the sphere. Consider the steady Bernoulli Equation~\ref{eqnBernoulliSteady} on a streamline infinitesimally close to the surface of the sphere. This gives:
\begin{equation*}
    \frac12 \rho \left[-\frac32 U \sin(\theta)\right]^2 + p = \frac12 \rho U^2 + p_{atm}
\end{equation*}
by using our previous results for $\vec{u}$.

On the sphere,
\begin{equation*}
    p - p_{atm} = \frac12 \rho U^2\left[1 - \frac94 \sin^2(\theta)\right]
\end{equation*}
and so the pressure distribution is symmetric up/down and front/back. Therefore, the net force on the sphere is zero!

This is an example of d'Alambert's Paradox which says that there is no drag (no force parallel to the motion of the fluid) on any steadily moving rigid body in an unbounded potential flow. This is the clearest example we have seen yet that ignoring the viscosity term is not a valid assumption in general. In real life, there will be a drag due to viscous effects.

However, the potential flow inviscid approximation is fine for bubbles, rapid acceleration of rigid particles and for small amplitude oscillations.
\subsection{Steady Flow on a Translating Cylinder}
We will again require a circulation $\kappa$. Consider the frame of the cylinder, and recall that the velocity potential is:
\begin{equation*}
    \phi = U\cos(\theta) \left(r + \frac{a^2}{r}\right) + \frac{\kappa}{2\pi}\theta
\end{equation*}
On $r = a$, $u_r = 0$ and $u_\theta = 2-U\sin(\theta) + \frac{\kappa}{2\pi}$.

By the Bernoulli Equation,
\begin{equation*}
    \frac12\rho \left[-2U\sin(\theta) + \frac{\kappa}{2\pi a}\right]^2 + p = \frac12 \rho U^2 + p_{atm}
\end{equation*}
\begin{align*}
    \vec{F} &= \text{force exerted (per unit length along its axis) on the cylinder} \\
    &= \int_S p \vec{n} dS \\
    &= -\int_{0}^{2\pi} \left[\frac12 \rho U^2 + p_{atm} - \frac12 \rho \left(4U^2 \sin^2{\theta} + \frac{\kappa^2}{4\pi^2 a^2} \right.\right. \\
    &\qquad\left.\left. - 2 U \sin{\theta} \frac{\kappa}{\pi a}\right)\right] \begin{pmatrix}\cos(\theta) \\ \sin(\theta)\end{pmatrix} a d\theta \\
    &= \begin{pmatrix}0 \\ -\rho U \kappa\end{pmatrix}
\end{align*}
\begin{remarks}
    \item Drag is still zero. D'Alambert's Paradox works in 2D and 3D.
    \item There is a new force perpendicular to the motion. This is the lift force. This force $F_y$ has the opposite sign to the circulation $\kappa$. We will see that this works for a general 2D shape in Ex3Q1.
\end{remarks}
\section{Unsteady Potential Flow}
\subsection{Free Oscillations in a Tube}
\begin{figure}
    \centering
    \begin{tikzpicture}[scale=1]
        % Draw the outer and inner boundaries of the U-tube
        \draw[very thick, rounded corners=1.7cm] (-2, 7) -- (-2, 0) -- (2, 0) -- (2, 7);
        \draw[very thick, rounded corners=1.1cm] (-1.4, 7) -- (-1.4, 0.6) -- (1.4, 0.6) -- (1.4, 7);

        % Draw the fluid levels
        \draw[very thick, color=blue!80!black] (-2, 2.5) -- (-1.4, 2.5);
        \draw[very thick, color=blue!80!black] (1.4, 5.5) -- (2, 5.5);

        % Draw the dashed equilibrium line
        \draw[dashed, thick] (-2.3, 4) -- (2.3, 4);

        % Draw the bottom reference line
        \draw[thick] (-2.5, 0) -- (2.5, 0);

        % Add labels
        \node[left] at (-2.2, 2.5) {\Large $z=h-\zeta(t)$};
        \node[right] at (2.2, 5.5) {\Large $z=h+\zeta(t)$};
        \node[right] at (2.3, 4) {\Large $z=h$};
        \node[right] at (2.5, -0.15) {\Large $z=0$};
    \end{tikzpicture}
    \caption{Diagram of flow in a tube}
    \label{figFlowTube}
\end{figure}
Consider flow in a tube as in figure~\ref{figFlowTube}. Let the flow be initially irrotational (and so the flow will remain irrotational). Let the tube have uniform cross section, with radius $a$. Reduce the problem to two long tubes, with uniform flow in each tube. Note that $u = \dot{\zeta}$, so on the right we have $\phi = \dot{\zeta} z$ and on the left $\phi = -\dot{\zeta} z$. Consider \eqnref{eqnBernoulliUnsteady} evaluated at the top of the fluid on each side.
\begin{align*}
    \rho \left. \frac{\partial \phi}{\partial t}\right|_{h + \zeta} + \frac12 \rho \dot{\zeta}^2 + p_{atm} + \rho g (h + \zeta) &= \rho \left. \frac{\partial \phi}{\partial t}\right|_{h - \zeta} + \frac12 \rho \dot{\zeta}^2 + p_{atm} + \rho g (h - \zeta) \\
    \rho \ddot{\zeta} (h + \zeta) + \rho g (h + \zeta) &= -\rho \ddot{\zeta}(h - \zeta) + \rho g (h - \zeta) \\
    \rho \ddot{\zeta} h + \rho g \zeta &= -\rho \ddot{\zeta} h - \rho g \zeta \\
    2\rho h \ddot{\zeta} + 2 \rho g \zeta &= 0
\end{align*}
and so the equation of motion is $\ddot{\zeta} = \frac{g}{h} \zeta = 0$, which is oscillatory motion with frequency $\sqrt{\frac{g}{h}}$.
\subsection{Oscillation of a Bubble}
Consider a compressible bubble of gas of radius $a(t)$ in a liquid of density $\rho$. Assume spherical symmetry, so the flow must be radial, $u = \frac1{r^2}$. By ensuring $u(a) = \dot{a}$,
\begin{equation*}
    \vec{u} = \dot{a} \frac{a^2}{r^2} \vec{e_r}
\end{equation*}
and the velocity potential is:
\begin{equation*}
    \phi = -a^2 \frac{\dot{a}}{r}
\end{equation*}
In order to apply \eqnref{eqnBernoulliUnsteady}, we need:
\begin{align*}
    \frac{\partial \phi}{\partial t} &= -\frac{2\dot{a}^2 a + a \ddot{a}}{r} \\
    \left.\rho \frac{\partial \phi}{\partial t} \right|_{r = a} &= -\rho \left(2\dot{a}^2 + \ddot{a} a\right) \\
\end{align*}
and then \eqnref{eqnBernoulliUnsteady} becomes:
\begin{align*}
    \rho \frac{\partial \phi}{\partial t} &= \frac12 \rho |\vec{u}|^2 + p = f(t) \\
    -\frac32 \rho \dot{a}^2 - \rho a \ddot{a} &= p_{atm} - p(a, t)
\end{align*}
Now we require small oscillations, so $a = a_0 + \eta$ where $\eta \ll a_0$.
\begin{equation*}
    \rho a_0 \ddot{\eta} = p - p_{atm}
\end{equation*}
then we need to understand how $p$ changes with $a$. We will assume $PV^{\gamma} = \text{const}$. We linearise this to get:
\begin{equation*}
    \frac{\delta p}{p_{atm}} = -3\gamma \frac{\eta}{a_0}
\end{equation*}
and so,
\begin{equation*}
    \rho a_0 \ddot{\eta} = -3\gamma \frac{p_{atm}}{a_0} \eta
\end{equation*}
which is small oscillations with frequency
\begin{equation*}
    \omega = \sqrt{\frac{3\gamma p_{atm}}{\rho a_0^2}}
\end{equation*}
\subsection{Accelerating Sphere}
Consider a sphere accelerating through a fluid in the lab frame. Recall from section~\ref{secTranslatingSphere},
\begin{equation*}
    \phi = \frac{U\cos{\theta}a^3}{2r^2} \implies \rho = -\frac{\vec{U}(t) \cdot \vec{x}}{2r^3}a^3
\end{equation*}
and so,
\begin{align*}
    \rho \frac{\partial \phi}{\partial t} &= -\rho \frac{\dvec{U} \cdot \vec{x}}{2r^3}a^3 \\
    \implies \left.\rho \frac{\partial \phi}{\partial t} \right|_{r = a} &= \frac12 \rho \dvec{U} \cdot \vec{x}
\end{align*}
and so we can now use \eqnref{eqnBernoulliUnsteady} to find:
\begin{equation*}
    p = \underbrace{p_{atm} + \frac12 \rho U^2 (1 - \frac{9}{4}\sin^2{\theta})}_{\text{steady terms (section~\ref{secTranslatingSphere})}} + \underbrace{\frac12 \rho \dvec{U}\cdot\vec{x}}_{\text{new unsteady terms}} \\
\end{equation*}
The force on the sphere is then:
\begin{align*}
    \vec{F} &= -\int_S p \vec{dS} \\
    &= \underbrace{0}_{\text{steady part}} - \frac12 \rho \dvec{U} \cdot \int_S \vec{x} \vec{dS} \\
    &= -\frac12 \rho \dvec{U} \cdot \frac43 \pi a^2 I \text{ by isotropic tensor} \\
    &= -\frac12 \rho V \dvec{U} \\
    &= -m^* \dvec{U}
\end{align*}
where $m^*$ is the \underline{added mass}, proportional to the fluid displaced (where the constant of proportionality depends on the shape). To see very clearly why this is called the added mass, consider:
\begin{align*}
    \vec{F_{ext}} - m^* \dvec{U} &= m \dvec{U} \\
    \therefore (m + m^*) \dvec{U} &= \vec{F_{ext}}
\end{align*}
\end{document}