\documentclass[../Main.tex]{subfiles}

\begin{document}
\section{LU Factorisation}
\subsection{Motivation}
\begin{definition}{Unit lower triangular matrix}
    A matrix $M \in \R^{n \times n}$ is \underline{unit lower triangular} if it is lower triangular and all elements on the leading diagonal are $1$.
\end{definition}
\begin{definition}{LU factorisation}
    Let $A \in \R^{n \times n}$. We say that $A = L U$ is an \underline{LU factorisation} if $L \in \R^{n \times n}$ is unit lower triangular and $U \in \R^{n \times n}$ is upper triangular.
\end{definition}
Then to see an immediate use,
\begin{equation*}
    \det(A) = \underbrace{\det(L)}_{1} \det(U) = \prod_{i=1}^{n} U_{ii}
\end{equation*}
This has complexity $O(n)$, compared to the $O(n!)$ complexity of computing a determinant in general.

Throughout this chapter we will be interested in solving:
\begin{equation}
    A \vec{x} = \vec{b}
    \label{eqnLinearSystemEqs}
\end{equation}
where $A \in \R^{n \times n}$ and $\vec{x}, \vec{b} \in \R^n$. With an $LU$ factorisation, we solve instead $LU \vec{x} = \vec{b}$. Solving this is far simpler and we can break it into two very simple problems:
\begin{equation*}
    L\vec{y} = \vec{b} \text{ and } U\vec{x} = \vec{y}.
\end{equation*}
The first breaks down into a very simple process:
\begin{align*}
    y_1 &= b_1 \\
    y_2 &= b_2 - L_{2,1} y_1 \\
    &\vdots\\
    y_n = b_n - \sum_{j=1}^{n-1} L_{n,j} y_j
\end{align*}
and this is called forward substitution. At each step, we have an explicit equality in terms of variables that are already calculated. Once we find $\vec{y}$, we next solve $U\vec{x} = \vec{y}$:
\begin{align*}
    x_n &= \frac{1}{U_{n, n}} y_n \\
    x_{n-1} &= \frac{1}{U_{n-1, n-1}} (y_{n-1} - U_{n-1, n} x_n) \\
    &\vdots\\
    x_1 &= \frac{1}{U_{11}} \left(y_1 = \sum_{j=1}^{n-1} U_{1, j} x_j\right)
\end{align*}
and this is called backward substitution. The computational complexity for the forward substitution is:
\begin{equation*}
    2\sum_{j=1}^{n} j = n(n+1) = O(n^2)
\end{equation*}
and the same for the backward substitution, so the whole process has complexity $O(n^2)$.
\subsection{The Algorithm}
For some notation, let the $i$th row of a matrix $M$ be denoted by $\vec{m_i}^T$ and the $j$th column by $\vec{M_j}$. We will be interested in $L$ in terms of its columns $\vec{L_j}$ and $U$ in terms of its rows $\vec{u_i}^T$.

We can then think of the matrix $A$ as a sum of outer products of rows and columns.
\begin{equation*}
    LU = \sum_{i=1}^{n} \vec{L_i} \otimes \vec{u_i} = \vec{L_i} \vec{u_i}^T
\end{equation*}
Given the structure of $\vec{L_j}$ and $\vec{u_i}$, the matrix formed by this outer product has the block form:
\begin{equation*}
    \begin{pmatrix}
        0 & 0 \\
        0 & M
    \end{pmatrix}
\end{equation*}
where the first $(i-1)$ rows and columns are zero. We can take advantage of this, because the contribution of the first row and column can only come from $\vec{L_1} \vec{u_1}^T$. Therefore, we pick $\vec{L_1}$ to be the first column of $A$, divided by $A_{11}$ such that it has first element $1$. Pick $\vec{u_1}^T$ to be the first row of $A$. By using this choice, $LU$ will match on the first row and column with $\vec{L_1} \vec{u_1}^T$.

To find the next outer product, $\vec{L_2} \vec{u_2}^T$, subtract $\vec{L_1} \vec{u_1}^T$ from $A$ and apply the same process to row $2$ and column $2$ (this matrix will have row $1$ and column $1$ all zero, as required). Continue until $L$ and $U$ are determined. Then, our algorithm looks like:
\begin{enumerate}
    \item Input $A \in \R^{n \times n}$.
    \item Let $A_0 = A$.
    \item For $k = 1$ to $n$,
    \begin{enumerate}
        \item Take the $k$th column of $A_{k-1}$, and divide it by its first element. Let this be $\vec{L_k}$.
        \item Take the $k$th row of $A_{k-1}$, and let this be $\vec{u_k}^T$
        \item Set $A_k = A_{k-1} - \vec{L_k} \vec{u_k}^T$.
    \end{enumerate}
\end{enumerate}
\subsection{Computational Complexity}
The computational complexity comes from step 3c. The cost of finding $\vec{L_k} \vec{u_k}^T$ is $(n-k)^2$, and so the overall computational complexity from this is:
\begin{equation*}
    \sum_{k=1}^{n} (n-k)^2 = \sum_{i=0}^{n-1}i^2 = O(n^3)
\end{equation*}
This $O(n^3)$ shows up a lot in Numerical Linear Algebra. At its core, this is due to the problem of multiplying matrices. In order to compute $C = AB$, where $A, B \in \R^{n \times n}$, we compute:
\begin{equation*}
    c_{ij} = \sum_{k=1}^{n} A_{ik} B_{kj}
\end{equation*}
which is $O(n^3)$ for the whole matrix. We can find an obvious lower bound for the matrix multiplication, which must be $O(n^2)$. Indeed, there exists an algorithm with complexity $O(n^{\log_2(7)})$ which is Strassen's Algorithm. There exist algorithms with lower complexity, but the constant in front of the leading term is so large that these algorithms are not practical.
\subsection{Relation to Gaussian Elimination}
In Gaussian Elimination, we solve $A\vec{x} = \vec{b}$ by applying elementary row operations to $A$ and $\vec{b}$. We can understand the EROs as matrices $L_i$, which are unit lower triangular. Therefore we find:
\begin{equation*}
    L_{n-1} \cdots L_2 L_1 A = U
\end{equation*}
where $U$ is upper-triangular. Therefore we can find $L$ such that $A = LU$ by inverting the product $L_{n-1} \cdots L_2 L_1$. Since unit lower triangular matrices form a group under matrix multiplication, this inverse product will also be unit lower triangular.
\subsection{Dealing With Zero Entries}
So far, the algorithm we have written down cannot deal with if there is a zero diagonal entry. To resolve this, we use \underline{pivoting}. If we encounter a zero diagonal entry when performing the algorithm, we will permute the row with a row that has a nonzero element in the required column. In terms of the algorithm, this is:
\begin{enumerate}
    \item Input $A \in \R^{n \times n}$.
    \item Let $A_0 = A$.
    \item For $k = 1$ to $n$,
    \begin{enumerate}
        \item Find $p \in \{k, \cdots, n\}$ such that $(A_{k-1})_{p, k} \neq 0$.
        \item Let $P_k$ be the matrix corresponding to a transposition of the $k$th row with the $p$th.
        \item Now take the $k$th column of $P_k A_{k-1}$, and divide it by its first element (which we know is now nonzero). Let this be $\vec{L_k}$.
        \item Take the $k$th row of $P_kA_{k-1}$, and let this be $\vec{u_k}^T$
        \item Set $A_k = P_kA_{k-1} - \vec{L_k} \vec{u_k}^T$.
    \end{enumerate}
\end{enumerate}
\begin{remarks}
    \item There still exists an edge case, when we discover a column that is all zero. In this case, we could take $(l_k)_j = \delta_{j, k}$ and $\vec{u_k}^T$ as normal.
    \item In the case that a diagonal entry is very small, but nonzero, we might decide to apply pivoting to avoid very large numbers.
\end{remarks}
We \textit{unroll} the algorithm to understand what is happening:
\begin{align*}
    A_0 &= A \\
    A_1 &= P_1 A_0 - \vec{L_1} \vec{u_1}^T \\
    A_1 &= P_2 A_1 - \vec{L_2} \vec{u_2}^T = P_2 P_1 A - P_2 \vec{L_1} \vec{u_1}^T - \vec{L_2} \vec{u_2}^T \\
    &\vdots\\
    A_n &= \underbrace{P_n \cdots P_2 P_1}_{P} - \sum_{i=1}^{n} \tilde{\vec{L}}_i \vec{u_i}^T
\end{align*}
where $\tilde{\vec{L}}_i = \left(\prod_{j=i+1}^{n} P_j\right)\vec{L_i}$. Note that $A_n = 0$ by construction, and so:
\begin{equation*}
    PA = \sum_{i=1}^{n} \tilde{\vec{l}}_i \vec{u_i}^T = \tilde{L} U
\end{equation*}
\begin{remark}
    Not all matrices have an LU factorisation, but all matrices have a pivoted LU factorisation.
\end{remark}
\subsection{Special Cases of Matrices}
Consider first symmetric matrices, $A \in \R^{n \times n}$ where $A = A^T$. We have seen this in the Newton-Raphson method, where we need to solve a linear system of equations involving the Jacobian. In this case, we use a specific type of LU factorisation.
\begin{definition}{$LDL^T$ factorisation}
    A symmetric matrix $A$ has an \underline{$LDL^T$ factorisation} if there exists $L \in \R^{n \times n}$ that is unit lower triangular, and $D \in \R^{n \times n}$ which is diagonal, such that $A = LDL^T$.
\end{definition}
Clearly this is symmetric, so this factorisation can only exist for symmetric matrices. It is a special case of $LU$ factorisation where $U = DL^T$. The algorithm for this is similar to the algorithm for more general $LU$ factorisation. Let the columns of $L$ be $\vec{L_i}$. Then the matrix $LDL^T$ is an outer product:
\begin{equation*}
    A = \sum_{i=1}^{n} D_{ii} \vec{L_i} \otimes \vec{L_i}^T
\end{equation*}
and again we can use the fact that, for entry $i$ in this sum, rows and columns less than $i$ are zero.

\begin{enumerate}
    \item Input $A \in \R^{n \times n}$ such that $A = A^T$.
    \item Let $A_0 = A$
    \item For $k = 1$ to $n$,
    \begin{enumerate}
        \item Take the $k$th column of $A_{k-1}$ and divide it by its first element. Let this be $\vec{L_k}$.
        \item Let $D_{kk}$ be $(A_{k-1})_{kk}$.
        \item Set $A_k = A_{k-1} - D_{kk} \vec{L_k} \vec{L_k}^T$.
    \end{enumerate}
\end{enumerate}
\begin{remark}
    In order to deal with zero entries, we need to apply pivoting. However, to keep the symmetry we will also need to swap columns, resulting in a factorisation of $PAP^T$.
\end{remark}

\begin{definition}{Positive definiteness}
    A real symmetric matrix is \underline{positive definite} if $\vec{x}^T A \vec{x} > 0$ for all $\vec{x} \in \R^n \backslash \{\vec0\}$.
\end{definition}
\begin{remark}
    Equivalently, all eigenvalues of $A$ are positive.
\end{remark}
\begin{theorem}
    Let $A \in \R^{n \times n}$ be real symmetric. Then $A$ is positive definite if and only if $A$ has an $LDL^T$ factorisation with $D_{kk} > 0$.
    \label{thmLDLPosDef}
\end{theorem}
\begin{proof}
    \begin{proofdirection}{$\Leftarrow$}{Let $D_{kk} > 0$}
        For any $\vec{x} \in \R^n \backslash \{\vec0\}$,
        \begin{align*}
            \vec{x}^T A \vec{x} &= \vec{x}^T L D L^T \vec{x} \\
            &= \vec{y}^T D \vec{y} \text{ where $\vec{y} = L^T \vec{x} \neq \vec0$ because $L$ is unit lower triangular} \\
            &> 0 \text{ because $D$ has positive entries}
        \end{align*}
    \end{proofdirection}
    \begin{proofdirection}{$\Rightarrow$}{Let $A$ be positive definite}
        We need to show that $(A_{k-1})_{kk} > 0$, using the same notation as in the algorithm.

        Let $D_{kk} = (A_{k-1})_{kk} = \vec{x}^T A \vec{x}$ for some $\vec{x} \neq \vec0$. Next, require $x_k = 1$ and $x_i = 0$ for $i > k$. For such an $\vec{x}$,
        \begin{equation*}
            \vec{x}^T A_{k-1} \vec{x} = \sum_{i,j=1}^{n} (A_{k-1})_{ij} 
        \end{equation*}
        Recall that $A_{k-1} = A - \sum_{i=1}^{k-1} D_{ii} \vec{L_i} \vec{L_i}^T$. Therefore,
        \begin{equation*}
            D_{kk} = \vec{x}^T \left(A- \sum_{i=1}^{k-1} D_{ii} \vec{L_i} \vec{L_i}^T\right)\vec{x} = \vec{x}^T A\vec{x} - \sum_{i=1}^{k-1} D_{ii} \left(\vec{L_i}^T \vec{x}\right)^2
        \end{equation*}
        and so now we define $\vec{x}$ such that:
        \begin{equation*}
            \begin{cases}
                \vec{L_i}^T \vec{x} = 0 & \forall i \in \{1, \cdots, k-1\} \\
                x_k = 1 &  \\
                x_i = 0 & \forall i \in \{k+1, \cdots, n\}
            \end{cases}
        \end{equation*}
        Then the system defined by $\vec{L_i}^T \vec{x} = 0$ is given by $\tilde{L} \vec{x} = 0$ where $\tilde{L}$ is an upper triangular matrix. Therefore there must exist a solution in $\R \backslash \{0\}$. Now finally,
        \begin{align*}
            D_{kk} &= \vec{x}^T A_{k-1} \vec{x} \\
            &= \vec{x}^T A \vec{x} - \sum_{i=1}^{k-1} D_{ii} \underbrace{\left(\vec{L_i}^T \vec{x}\right)^2}_{0} \\
            &= \vec{x}^T A \vec{x} > 0
        \end{align*}
        as required.
    \end{proofdirection}
\end{proof}
\begin{remark}
    We can therefore use the $LDL^T$ factorisation to check if a matrix is positive-definite.
\end{remark}
\begin{example}[Cholesky Factorisation]
    If $A$ is positive-definite and $A = LDL^T$, where $D$ has positive diagonal entries, we can write:
    \begin{equation*}
        A = L\sqrt{D} \left(\sqrt{D}\right)^T L^T = \tilde{L} \tilde{L}^T
    \end{equation*}
    where $\sqrt{D}$ is the matrix with diagonal entries $(\sqrt{D})_{ii} = \sqrt{D_{ii}}$, and $\tilde{L} = L\sqrt{D}$.

    Note that now $\tilde{L}$ is lower triangular but not unit lower triangular. This factorisation, $A = \tilde{L} \tilde{L}^T$, is called the Cholesky Factorisation.
\end{example}
\subsection{Sparsity}
We are interested in matrices with many of their elements zero. Such matrices are called \underline{sparse}.

\begin{definition}{Banded matrix}
    A \underline{banded matrix} is a matrix with non-zero elements only near the leading diagonal. That is, for some $1 \leq r < n-1$, the matrix has $a_{ij} = 0$ if $|i - j| > r$. This value $r$ is called the \underline{band width}.
\end{definition}
Now we can give the following theorem to understand the LU factorisation of a banded matrix.
\begin{theorem}
    Let $A$ be a matrix with an $LU$ factorisation $A = LU$. Assume that the diagonal elements are non-zero in $A$. Then:
    \begin{enumerate}
        \item All leading zeroes in any row of $A$ to the left of the diagonal are inherited by $L$.
        \item All leading zeroes in any column of $A$ above the diagonal are inherited by $U$.
    \end{enumerate}
    \label{thmSparseLU}
\end{theorem}
\begin{remark}
    This statement is only about leading zeroes, so for a row in $A$ of the form $(0, 0, 0, 4, 0, 0, 0)$, only the first $3$ zeroes are inherited, by which we mean that the first $3$ entries of the corresponding row in $L$ are zero.
\end{remark}
\begin{proof}
    For part 1, let $i \in \{1, \cdots, n\}$ be the index of a row, and $A_{i, 1} = \cdots = A_{i,q} = 0$, where $q < i$.
    \begin{align*}
        0 &= A_{i,1} (LU)_{i,1} = L_{i,1} U_{1, 1} \\
        \implies L_{i,1} &= 0 \text{ because $A$ has non-zero diagonal entries} \\
        0 &= A_{i,2} = (LU)_{i,2} = L_{i,1} U_{1,2} + L_{i,2} U_{2,2} \\
        \implies L_{i,2} &= 0 \text{ since $L_{i,1} = 0$, $U_{2, 2} \neq 0$}
    \end{align*}
    then we can continue this argument until $L_{i,q}$.

    The argument for part 2 (about $U$) is the same.
\end{proof}
Now, for a banded matrix of band width $r$, we know that only the $r-1$ entries below the diagonal will be non-zero in $L$ (with all $1$s on the diagonal), and only the diagonal elements and first $r-1$ entries to the right will be non-zero in $U$. In this way, we see that the banded structure is inherited by $L$ and $U$. Further, the complexity of the $LU$ factorisation algorithm for a banded matrix of band width $r$ is $O(r^2n)$.
\begin{example}[Sparse matrix with dense $LU$ factorisation]
    Consider a matrix with non-zero elements in the first column, first row and leading diagonal (an arrow matrix). In this case, \thmref{thmSparseLU} fails, and $L$ and $U$ will be dense.

    We can perform a row swap and a column swap of the first and last (pivoting) to create a reverse arrow matrix, in which the non-zero elements are in the first column, last column and leading diagonal only. Then \thmref{thmSparseLU} gives that $L$ and $U$ are very sparse.
\end{example}
\section{QR Factorisation}
\begin{definition}{Orthogonal matrix}
    A matrix $Q$ is \underline{orthogonal} if $QQ^T = Q^TQ = I$.
\end{definition}
From the course IA Vectors and Matrices or IB Linear Algebra, we know that, equivalently, the columns of $Q$ form an orthonormal basis under a given inner product.
\begin{definition}{QR factorisation}
    Let $A \in \R^{m \times n}$. We say that $A = QR$ is a \underline{(full) QR factorisation} of $A$ if:
    \begin{enumerate}
        \item $Q \in \R^{m \times m}$ is an orthogonal matrix,
        \item $R \in \R^{m \times n}$ is an upper triangular matrix ($R_{ij} = 0$ if $i > j$).
    \end{enumerate}
\end{definition}
\begin{definition}{Reduced QR factorisation}
    For $A \in R^{m \times n}$, with $m > n$, a \underline{reduced QR factorisation} is $A = QR$ where:
    \begin{enumerate}
        \item $Q \in \R^{m \times n}$ has orthonormal columns ($Q^TQ = I$),
        \item $R \in \R^{n \times n}$ is upper triangular.
    \end{enumerate}
\end{definition}
We see the relevance of the reduced QR factorisation:
\begin{equation*}
    \begin{bNiceMatrix}[first-row,first-col]
        &&n&\\
        &\Block{3-3}{A}&&\\
        m&&&\\
        &&&
    \end{bNiceMatrix}
    =
    \begin{bNiceMatrix}[first-row,last-row,vlines=3]
        &n&\hphantom{n}&\hphantom{nnn}\\
        &\Block{3-1}{Q_1}&\Block{3-2}{Q_2}&\\
        &&&\\
        &&&\\
        &&Q&
    \end{bNiceMatrix}
    \begin{bNiceMatrix}[last-row,hlines=3]
        \Block{2-2}{\diagbox{~0}{R_1}}\\
        &\\
        \Block{1-2}{0}\\
        R
    \end{bNiceMatrix}
    = Q_1 R_1 + (0)
\end{equation*}
where $Q_1 R_1$ is the reduced $QR$ factorisation. We can create $Q$ from $Q_1$ by choosing any set of columns that are orthonormal to themselves and the columns of $Q_1$ (any such choice is valid because we then multiply them by zero).
\subsection{Application to Solving Linear Systems}
Consider:
\begin{equation*}
    A\vec{x} = \vec{b} \tag{\ref{eqnLinearSystemEqs}}
\end{equation*}
then we solve $QR\vec{x} = \vec{b}$ which is equivalent to $R\vec{x} = Q^T\vec{b}$ and since $R$ is upper-triangular, we can use backward substitution. Both the product $Q^T\vec{x}$ and backward substitution have complexity $O(n^2)$, so solving $QR\vec{x}$ has complexity $O(n^2)$.
\subsection{Gram-Schmidt Algorithm}
If we consider $A$ in terms of its columns $\vec{A_j}$, and $Q$ in terms of its (orthonormal) columns $\vec{Q_j}$, we see that:
\begin{equation*}
    \vec{A_i} = \sum_{i=1}^{j} R_{ij} \vec{Q_j}
\end{equation*}
and we see clearly that Gram-Schmidt is equivalent to this process. The classical Gram-Schmidt algorithm is:
\begin{enumerate}
    \item For $i = 1$ to $n$,
    \begin{enumerate}
        \item let $\vec{b_i} = \vec{a_i} - \sum_{i = 1}^{j-1} \frac{\inn{\vec{a_j}}{\vec{b_i}}}{\norm{\vec{b_i}}^2}\vec{b_i}$.
        \item Let $\vec{Q_i} = \frac{\vec{b_i}}{\norm{\vec{b_i}}}$.
        \item Let $R_{ij} = \inn{\vec{A_j}}{\vec{q_i}}$ for $i < j$,
        \item let $R_{ii} = \norm{\vec{b_i}}$
    \end{enumerate}
    \item Let $Q$ have columns $\vec{Q_i}$, $R$ have elements $R_{ij}$.
\end{enumerate}
\begin{remarks}
    \item This algorithm computes the reduced $QR$ factorisation.
    \item The total complexity is $O(mn^2)$.
    \item Gram-Schmidt is rarely used in practice, because it does not behave well when there exists a small error in calculations (machine precision - around $10^{-16}$), and the error is multiplied during the calculation. The next algorithms behave much better with this error.
    \item We can describe this algorithm as ``triangular orthogonalisation'', applying triangular operations to orthogonalise $Q$. The alternative is ``orthogonal triangularisation'', where orthogonal operations are applied to $R$ to make it triangular. These will be seen in the next sections.
\end{remarks}
\section{Givens Rotations}
\subsection{Givens Rotation Matrices}
\begin{definition}{Givens rotation matrix}
    A \underline{Givens rotation matrix} is an $n \times n$ matrix $\Omega^{[p,q]}(\theta)$, where $1 \leq p < q \leq n, \theta \in [0, 2\pi)$. It has the form:
    \begin{equation*}
        \begin{bNiceMatrix}[first-row,first-col]
            &&&& p &&&& q &&&\\
            &1&&&&&&&&&&\\
            &&\Ddots&&&&&&&&&\\
            &&&1&&&&&&&&\\
            p &&&& \cos(\theta) &&&& \sin(\theta) &&&\\
            &&&&&1&&&&&&\\
            &&&&&&\Ddots&&&&&\\
            &&&&&&&1&&&&\\
            q &&&& -\sin(\theta) &&&& \cos(\theta) &&&\\
            &&&&&&&&&1&&\\
            &&&&&&&&&&\Ddots&\\
            &&&&&&&&&&&1\\
        \end{bNiceMatrix}
    \end{equation*}
\end{definition}
The intuition behind the algorithm is as follows. We want to left-apply Givens rotation matrices to the matrix $A$ in order to set the elements below the diagonal to zero. However, we must be careful in later steps to preserve the zeroes we already have.
\begin{theorem}
    Let $A \in \R^{m \times n}$. Let $i \in \{1, \cdots, m\}$ and $j \in {1, \cdots, n}$ with $i > j$.

    There exists a Givens rotation matrix $\Omega^{[p,q]}$ with $p < i$ and $q = i$ such that:
    \begin{equation*}
        \left(\Omega^{[p,q]}A\right)_{ij} = 0.
    \end{equation*}
    Furthermore, all the rows of $\Omega^{[p,q]}A$ other than $p$ and $q$ are equal to those of $A$. Rows $p$ and $q$ in $\Omega^{[p,q]}$ are a linear combination of rows $p$ and $q$ of $A$.
    \label{thmGivensEffect}
\end{theorem}
\begin{proof}
    We take:
    \begin{equation*}
        \cos(\theta) = \frac{A_{pj}}{\sqrt{A_{pj}^2 + A_{qj}^2}},\qquad \sin(\theta) = \frac{A_{qj}}{\sqrt{A_{pj}^2 + A_{qj}^2}}
    \end{equation*}
    which satisfies the conditions.
\end{proof}
\subsection{The Algorithm}
\begin{enumerate}
    \item Let $Q = I$.
    \item For $j = 1$ to $n$, for $i = j+1$ to $m$,
    \begin{enumerate}
        \item Find $\Omega^{[j,i]}$ such that $\left(\Omega^{[j,i]}A\right)_{ij} = 0$ (as in \thmref{thmGivensEffect}).
        \item Overwrite $A$ with $\Omega^{[j,i]}A$.
        \item Overwrite $Q$ with $Q\left(\Omega^{[i,j]}\right)^T$
    \end{enumerate}
    \item At the end of the algorithm, $A$ is upper-triangular and we re-label it to $R$.
\end{enumerate}
\begin{remark}
    The complexity of this algorithm is $O(mn^2)$.
\end{remark}
\section{Householder Reflections}
\begin{definition}{Householder reflection matrix}
    A \underline{Householder reflection matrix} $H = H_{\vec{u}} \in \R^{m \times m}$ is a matrix that corresponds to a reflection in the $(m-1)$-surface normal to the non-zero vector $\vec{u}$:
    \begin{equation*}
        H = I - 2 \frac{\vec{u} \vec{u}^T}{\norm{\vec{u}}^2}
    \end{equation*}
\end{definition}

In this algorithm, we choose our reflection matrix such that the lower-diagonal elements of column $i$, $\vec{A_i}$, is mapped to a basis vector (and so has all zeroes below the diagonal). We therefore choose a hyperplane normal to $\vec{A_i} - \norm{\vec{A_i}}\vec{e_j}$. Note that this is non-unique, and depends on the target basis vector $\vec{e_j}$.

The algorithm is as follows:
\begin{enumerate}
    \item Input $A \in \R^{m \times n}$.
    \item For $j = 1$ to $n$,
    \begin{enumerate}
        \item Let $\vec{a} \in \R^{m - j + 1}$ be the elements of the $j$th column on and below the diagonal.
        \item Let $\vec{e}$ be the first basis vector of $\R^{m - j + 1}$.
        \item Define $\vec{u} = \vec{a} \pm \norm{\vec{a}}\vec{e}$
        \item Overwrite $A$ with:
            \begin{equation*}
                \begin{bmatrix}
                    I_{j - 1} & 0 \\
                    0 & H_{\vec{u}}
                \end{bmatrix}
                A
            \end{equation*}
    \end{enumerate}
\end{enumerate}
\begin{remarks}
    \item When updating $A$ in the above algorithm, we only need to consider the last rows that are affected by the Householder reflection submatrix. The other columns and rows are unchanged.
\end{remarks}
\end{document}