\documentclass[../Main.tex]{subfiles}

\begin{document}
\section{Definitions}
\begin{definition}{Hypothesis}
    A \underline{hypothesis} is an assumption about the distribution of data $\vec{X}$ taking values in $\samplespace$.
\end{definition}
\begin{definition}{Null hypothesis}
    The \underline{null hypothesis}, $H_0$ is the base case or ``no effect'' hypothesis.
\end{definition}
\begin{definition}{Alternative hypothesis}
    The \underline{alternative hypothesis}, $H_1$ is the hypothesis of a positive/negative effect.
\end{definition}
\begin{example}
    Consider flipping coins, $X_i \sim \text{Ber}(\theta)$. The null hypothesis is $H_0 : \theta = \frac12$. Examples of alternative hypotheses could be:
    \begin{itemize}
        \item $\theta = \frac34$
        \item $\theta \neq \frac12$
        \item $\theta < \frac12$
    \end{itemize}
\end{example}
\begin{example}[Goodness of Fit Test]
    Suppose that $X_i$ are independent and identically distributed random variable for $i = 1$ to $n$.
    We can then make hypotheses about the distribution function:
    \begin{align*}
        H_0:&X_i \text{ have pdf } f_1 \\
        H_1:&X_i \text{ have pdf } f_2 \\
    \end{align*}
\end{example}
\begin{example}
    Suppose that $\vec{X}$ has a pdf $f_{\vec{X}}(\vec{x}, \theta)$ for some $\theta \in \Theta$. We can make hypotheses about $\theta$:
    \begin{align*}
        H_0:&\theta \in \Theta_0 \subset \Theta \\
        H_1:&\theta \notin \Theta_0
    \end{align*}
\end{example}
\begin{definition}{Simple hypothesis}
    A hypothesis is \underline{simple} if it fully specifies the distribution of the data $\vec{X}$.

    The opposite of a simple hypothesis is a \underline{composite hypothesis}.
\end{definition}
\begin{definition}{Hypothesis test}
    A \underline{test} of a hypothesis $H_0$ is defined by a \underline{critical region} $C$. When $\vec{X} \in C$, we reject $H_0$. When $\vec{X} \notin C$, we do not reject $H_0$.
\end{definition}
\begin{remark}
    Note that when $\vec{X} \notin C$, we do not \textit{accept} $H_0$. We are simply saying that there is not enough data to reject it.
\end{remark}
\begin{definition}{Type I/II errors}
    A \underline{type I error} is the case of rejecting $H_0$ when $H_0$ was in fact true.
    A \underline{type II error} is the case of failing to reject $H_0$ when $H_1$ was true.
\end{definition}
\begin{definition}{Power/size}
    When $H_0$ and $H_1$ are simple hypotheses, the \underline{size} of the test is the probability of a type I error,
    \begin{equation*}
        \alpha = \P_{H_0}(H_0 \text{ rejected}) = \P_{H_0}(\vec{X} \in C).
    \end{equation*}
    The \underline{power} of the test is the probability of no type II errors. The power is $1 - \beta$ where:
    \begin{equation*}
        \beta = \P_{H_1}(H_0 \text{ not rejected}) = \P_{H_1}(\vec{X} \notin C)
    \end{equation*}
\end{definition}
Usually, we set $\alpha$ at an acceptable level and choose a test that minimises $\beta$ subject to the condition on $\alpha$.
\section{Neyman-Pearson Lemma}
\begin{definition}{Likelihood ratio statistic}
    Let $H_0$ and $H_1$ be simple, with $\vec{X}$ having pdf $f_i$ under $H_i$ ($i = 0$ or $1$). The \underline{likelihood ratio statistic} is:
    \begin{equation*}
        \Lambda_{\vec{X}}(H_0, H_1) = \frac{f_1(\vec{X})}{f_0(\vec{X})}
    \end{equation*}
\end{definition}
\begin{definition}{Likelihood ratio test}
    A \underline{likelihood ratio test} rejects $H_0$ when $\vec{X} \in C = \subsetselect{\vec{x}}{\Lambda_{\vec{X}}(H_0, H_1) > k}$.
\end{definition}
\begin{theorem}[Neyman-Pearson Lemma]
    Suppose $f_0$ and $f_1$ are non-zero on the same set and there exists $k$ such that the critical region:
    \begin{equation*}
        C_k = \subsetselect{\vec{x}}{\frac{f_1(\vec{x})}{f_0(\vec{x})} > k}
    \end{equation*}
    has size $\alpha$. Then for all tests with size $\leq \alpha$, the LRT is the test with smallest $\beta$.
    \label{thmNeymanPearson}
\end{theorem}
\begin{proof}
    Let $\overline{C}$ be the complement of $C$.
    \begin{align*}
        \alpha &= \P_{H_0}(\vec{X} \in C) = \int_C f_0(\vec{x}) d^n \vec{x} \\
        \beta &= \P_{H_1}(\vec{X} \in \overline{C}) = \int_{\overline{C}} f_1(\vec{x}) d^n \vec{x}
    \end{align*}
    Then let $C^\star$ be the critical region of another test of size $\alpha^\star \leq \alpha$. We want to show that $\beta \leq \beta^\star$.
    \begin{align*}
        \beta - \beta^\star &= \int_{\overline{C}} f_1(\vec{x}) d^n \vec{x} - \int_{\overline{C^\star}} f_1(\vec{x}) d^n \vec{x}\\
        &= \int_{\overline{C} \cap C^\star} f_1(\vec{x}) d^n\vec{x} - \int_{\overline{C^\star} \cap C} f_1(\vec{x}) d^n\vec{x} \\
        &= \int_{\overline{C} \cap C^\star} \underbrace{\frac{f_1(\vec{x})}{f_0(\vec{x})}}_{\text{$\leq k$ on $\overline{C}$}} f_0(\vec{x}) d^n\vec{x} - \int_{\overline{C^\star} \cap C} \underbrace{\frac{f_1(\vec{x})}{f_0(\vec{x})}}_{\text{$> k$ on $C$}} f_0(\vec{x}) d^n\vec{x} \\
        &\leq k\left(\int_{\overline{C} \cap C^\star} f_0(\vec{x}) d^n\vec{x} - \int_{\overline{C^\star} \cap C} f_0(\vec{x}) d^n\vec{x}\right) \\
        &= k\left(\int_{\overline{C} \cap C^\star} f_0(\vec{x}) d^n\vec{x} + \int_{C \cap C^\star} f_0(\vec{x}) d^n\vec{x} - \int_{C \cap C^\star} f_0(\vec{x}) d^n\vec{x}\right.\\
        &\qquad\left. - \int_{\overline{C^\star} \cap C} f_0(\vec{x}) d^n\vec{x}\right) \\
        &= k\left(\int_{C^\star} f_0(\vec{x}) d^n\vec{x} - \int_C f_0(\vec{x}) d^n\vec{x}\right) \\
        &= k(\alpha^\star - \alpha)
    \end{align*}
    Therefore, since we define $\alpha^\star \leq \alpha$, we find $\beta^\star \geq \beta$.
\end{proof}
\begin{remark}
    A likelihood ratio test with size $\alpha$ does not always exist for any given $\alpha$. However, we can always define a ``randomised'' test with exact size $\alpha$. That is, we define a region where $H_0$ is always rejected, and define a region where there is a certain probability that $H_0$ is rejected.
\end{remark}
\begin{example}
    Suppose $X_i \sim N(\mu, \sigma_0^2)$ for $i = 1$ to $n$. Let $\sigma_0^2$ be fixed and known. Consider the hypotheses:
    \begin{align*}
        H_0: & \mu = \mu_0 \\
        H_1: & \mu = \mu_1
    \end{align*}
    for fixed $\mu_1 > \mu_0$. The likelihood ratio is:
    \begin{align*}
        \Lambda_{\vec{X}}(H_0, H_1) &= \frac{(2\pi\sigma_0^2)^{-n / 2} \exp\left(-\frac{1}{2\sigma_0^2} \sum_{i=1}^{n} (X_i - \mu_1)^2\right)}{(2\pi\sigma_0^2)^{-n / 2} \exp\left(-\frac{1}{2\sigma_0^2} \sum_{i=1}^{n} (X_i - \mu_0)^2\right)} \\
        &= \exp\left(\frac{\mu_1 - \mu_0}{\sigma_0^2} n \overline{X} + \frac{n(\mu_0^2 - \mu_1^2)}{2\sigma_0^2}\right)
    \end{align*}
    then since $\Lambda_{\vec{X}}(H_0, H_1)$ is monotone in $\overline{X}$,
    \begin{equation*}
        \Lambda_{\vec{X}}(H_0, H_1) > k \iff \overline{X} > c
    \end{equation*}
    By the same logic, define $Z = \frac{\sqrt{n}(\overline{X} - \mu_0)}{\sigma_0}$ which is standard normal under $\mu_0$. Therefore, the rejection region is $\{Z > c'\}$ Therefore, with size $\alpha$, we choose:
    \begin{equation*}
        C = \subsetselect{\vec{X}}{Z > \Phi^{-1}(1-\alpha)}
    \end{equation*}
\end{example}
\section{\texorpdfstring{$p$}{p}-Values}
\begin{definition}{$p$-value}
    For a test with critical region of the form $\subsetselect{\vec{x}}{T(\vec{x}) > k}$, where $T$ is some statistic, the \underline{$p$-value} is:
    \begin{equation*}
        p = \P_{H_0}(T(\vec{X}) > T(\vec{x}^*))
    \end{equation*}
    where $\vec{x}^*$ is the observed data.
\end{definition}
\begin{remark}
    This is the probability of observing more extreme data under $H_0$.
\end{remark}
\begin{proposition}
    Under $H_0$, $p \sim U[0, 1]$.
    \label{propPValUnif}
\end{proposition}
\begin{proof}
    Let $F$ be the probability distribution function of $T$. Then:
    \begin{align*}
        \P_{H_0}(p < u) &= \P_{H_0}(1-F(T(\vec{X})) < u) \\
        &= \P_{H_0}(F(T(\vec{X})) > 1-u) \\
        &= \P_{H_0}(T(\vec{X}) > F^{-1}(1-u)) \\
        &= 1 - F(F^{-1}(1-u)) = u.
    \end{align*}
\end{proof}
\section{Tests and Confidence Intervals}
\begin{definition}{Acceptance region}
    The \underline{acceptance region} $A$ of a test is the complement of the critical region.
\end{definition}
\begin{theorem}
    Let $\vec{X}$ have pdf $f_{\vec{X}}(\vec{x}, \theta)$ for some $\theta \in \Theta$.
    \begin{enumerate}
        \item Suppose that for each $\theta_0 \in \Theta$, there exists a test of $H_0: \theta = \theta_0$ of size $\alpha$ with acceptance region $A(\theta_0)$. Then the set
            \begin{equation*}
                I(\vec{X}) = \subsetselect{\theta}{\vec{X} \in A(\theta)}
            \end{equation*}
            is a $100(1-\alpha)\%$ confidence set.
        \item Suppose that $I(\vec{X})$ is a $100(1-\alpha)$ confidence set for $\theta$. Then:
            \begin{equation*}
                A(\theta_0) = \subsetselect{\vec{x}}{\theta_0 \in I(\vec{x})}
            \end{equation*}
            is the acceptance region of a test of size $\alpha$ for $H_0 : \theta = \theta_0$.
    \end{enumerate}
    \label{thmTestsCIs}
\end{theorem}
\begin{proof}
    In both cases, $\theta_0 \in I(\vec{X}) \iff \vec{X} \in A(\theta_0)$.

    \begin{subproof}{$\P_{\theta_0}(\theta_0 \in I(\vec{X})) = 1-\alpha$}
        Using the equivalence, the LHS is:
        \begin{equation*}
            \P_{\theta_0}(\vec{X} \in A(\theta_0)) = \P_{\theta_0}(\text{do not reject $H_0$}) = 1-\alpha
        \end{equation*}
    \end{subproof}
    \begin{subproof}{$\P_{\theta_0}(\vec{X} \notin A(\theta_0)) = \alpha$}
        Using the equivalence again,
        \begin{equation*}
            \P_{\theta_0}(\vec{X} \notin A(\theta_0)) = \P_{\theta_0}(\theta_0 \notin I(\vec{X})) = \alpha
        \end{equation*}
    \end{subproof}
\end{proof}
\begin{example}
    Suppose $X_1, \cdots, X_n \sim N(\mu, \sigma_0^2)$ where $\sigma_0$ is known. We know a $100(1-\alpha)\%$ confidence interval for $\mu$:
    \begin{equation*}
        I(\vec{X}) = \overline{X} \pm \frac{\sigma_0^2}{\sqrt{n}} Z_{\frac{\alpha}{2}}
    \end{equation*}
    therefore we can use the second part of \thmref{thmTestsCIs} to construct a test of size $\alpha$ for $H_0: \mu = \mu_0$.
    \begin{equation*}
        A(\mu_0) = \subsetselect{\vec{x}}{\mu_0 \in I(\vec{x})}
    \end{equation*}
    and we can write instead that we reject $H_0$ when:
    \begin{equation*}
        \frac{\sqrt{n}|\mu_0-\overline{X}|}{\sigma_0} > z_{\alpha / 2}
    \end{equation*}
    Then this is a 2-sided LRT.
\end{example}
\section{The Chi-Squared Distribution}
Let $\vec{X}$ be a vector of $n$ independent standard normal random variables. We want to consider:
\begin{equation*}
    Y = \sum_{i=1}^{n} X_i^2 = |\vec{X}|^2
\end{equation*}
and we find that the distribution is:
\begin{equation*}
    f_Y(y) = \frac{1}{2^{n/2} \Gamma(n / 2)} y^{\frac{k}{2} - 1} e^{-y / 2}
\end{equation*}
for $y > 0$, $0$ otherwise.

\begin{definition}{Chi-Squared Distribution}
    We say that $Y$ has a \underline{Chi-squared distribution} with $n$ \underline{degrees of freedom} if $Y = |\vec{X}|^2$ (with $\vec{X}$ defined as above). We write $Y \sim \chi_n^2$.
\end{definition}
The $\chi^2$ distribution is a special case of the Gamma distribution:
\begin{equation*}
    \chi_n^2 = \Gamma\left(\frac{k}{2}, \frac12\right)
\end{equation*}
from this fact we deduce:
\begin{equation*}
    \E[Y] = k \qquad \Var(Y) = 2k
\end{equation*}
\section{Composite Hypotheses}
Previously we have considered $H_0$ and $H_1$ simple hypotheses, with error probabilities:
\begin{equation*}
    \alpha \in \P_{H_0}(\vec{X} \in C),\qquad \beta = \P_{H_1}(\vec{X} \notin C)
\end{equation*}
We will now consider $\vec{X}$ having probability density $f_{\vec{X}}(\vec{x}, \theta)$ and:
\begin{align*}
    H_0:& \theta \in \Theta_0 \subseteq \Theta \\
    H_1:& \theta \in \Theta_1 \subseteq \Theta 
\end{align*}
\begin{definition}{Power function}
    The \underline{power function} is defined to be $W(\theta) = \P_{\theta}(\vec{X} \in C)$.
\end{definition}
\begin{definition}{Size (composite hypotheses)}
    The \underline{size} of a test with composite hypotheses as above is the worst-case type I error:
    \begin{equation*}
        \alpha = \sup_{\theta \in \Theta_0} W(\theta)
    \end{equation*}
\end{definition}
\subsection{Uniformly Most Powerful Tests}
\begin{definition}{Uniformly most powerful tests}
    A test of size $\alpha$ of $H_0$ against $H_1$ is \underline{uniformly most powerful} (UMP) if:
    \begin{enumerate}
        \item $\sup_{\theta \in \Theta_0} W(\theta) \leq \alpha$
        \item For any other test of size $\alpha$, with power function $W^*$, we have $W(\theta) \geq W^*(\theta)$ for all $\theta \in \Theta_1$
    \end{enumerate}
\end{definition}
\begin{remark}
    Such a test may not exist. A good candidate, however, is an LRT.
\end{remark}
\begin{example}[One-sided test for normal location]
    Suppose $X_1, \cdots, X_n \sum N(\mu, \sigma_0^2)$ with $\sigma_0^2$ known. Consider the hypotheses:
    \begin{align*}
        H_0:&\mu \leq \mu_0 \\
        H_1:&\mu > \mu_0
    \end{align*}
    Recall that for the simple hypotheses, $H_0' : \mu = \mu_0$ and $H_1' : \mu = \mu_1$, the LRT had critical region:
    \begin{equation*}
        C = \subsetselect{\vec{x}}{z = \frac{\sqrt{n}(\overline{X} - \mu_0)}{\sigma_0}}
    \end{equation*}
    We will show that the same test is UMP for our composite hypotheses. Consider the power function:
    \begin{align*}
        W(\mu) &= \P_{\mu}(\text{reject $H_0$}) \\
        &= \P_\mu\left(\frac{\sqrt{n}(\overline{X} - \mu_0)}{\sigma_0} > z_\alpha\right) \\
        &= \P_\mu\left(\frac{\sqrt{n}(\overline{X} - \mu)}{\sigma_0} > z_\alpha + \frac{\sqrt{n}(\mu_0 - \mu)}{\sigma_0}\right) \\
        &= 1 - \Phi\left(z_\alpha + \frac{\sqrt{n}(\mu_0 - \mu)}{\sigma_0}\right)
    \end{align*}
    Then this is monotonically increasing, and it is $\alpha$ at $\mu = \mu_0$. Therefore, $\sum_{\mu \in \Theta_0} W(\mu) = \alpha$ and the test has size $\alpha$.

    Next, consider any other test with size $\leq \alpha$ with power function $W^*$. This must have size $\leq \alpha$ for $H_0'$ against $H_1$. This is because:
    \begin{equation*}
        W^*(\mu_0) \leq \sup_{\mu \in \Theta} \leq \alpha.
    \end{equation*}
    Thus, by Neyman-Pearson, $W(\mu_1) \geq W^*(\mu_1)$. Then since this argument works for any $\mu_1 > \mu_0$, we find:
    \begin{equation*}
        W(\mu) \geq W^*(\mu) ~\forall \mu \in \Theta_1
    \end{equation*}
\end{example}
\subsection{Generalised Likelihood Ratio Tests}
The generalised likelihood ratio statistic is:
\begin{equation*}
    \Lambda_{\vec{X}}(H_0, H_1) = \frac{\sup_{\theta \in \Theta_1}f_{\vec{X}}(\vec{x}, \theta)}{\sup_{\theta \in \Theta_0}f_{\vec{X}}(\vec{x}, \theta)}
\end{equation*}
and the generalised likelihood ratio test (GLRT) rejects $H_0$ when $\Lambda_{\vec{X}}$ is large.
\begin{example}[Two-sided test for normal location]
    Consdier $X_1, \cdots, X_n \sim N(\mu, \sigma_0^2)$ where $\sigma_0^2$ is known. We would like to test:
    \begin{align*}
        H_0:&\mu = \mu_0 \\
        H_1:&\mu \neq \mu_0
    \end{align*}
    Then $\Theta_0 = \{\mu_0\}$ and $\Theta_1 = \R \backslash \{\mu_0\}$. Next, the generalised likelihood ratio statistic is:
    \begin{align*}
        \Lambda_{\vec{X}}(H_0, H_1) &= \frac{(2\pi \sigma_0^2)^{n / 2} \exp\left(\frac{1}{2\sigma_0^2} \sum_{i=1}^{n} (X_i - \overline{X})^2\right)}{(2\pi \sigma_0^2)^{n / 2} \exp\left(\frac{1}{2\sigma_0^2} \sum_{i=1}^{n} (X_i - \mu_0)^2\right)} \\
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) &= \frac{1}{\sigma_0^2} \left(\sum_{i=1}^{n} (X_i - \mu_0)^2 - \sum_{i=1}^{n} (X_i - \overline{X})^2\right) \\
        &= \frac{n}{\sigma_0^2} (\overline(X) - \mu_0)^2
    \end{align*}
    Then because $\log$ is an increasing function, the GLRT rejects $H_0$ if $\frac{\sqrt{n}|\overline(X) - \mu_0|}{\sigma_0}$ is large. Under $H_0$, this has distribution $N(0, 1)$. Therefore the test that rejects $H_0$ if the given ratio is greater than $\Phi^{-1}(\alpha / 2)$ has size $\alpha$.

    Note that $2\log(\Lambda_{\vec{X}}(H_0, H_1)) \sim \chi_1^2$. Therefore, the critical region can also be written:
    \begin{equation*}
        \left\{x~\left|~\frac{n(\overline{X} - \mu_0^2)}{\sigma_0^2} > \chi_1^2(\alpha)\right.\right\}
    \end{equation*}
    \label{expTwoSideNormalTest}
\end{example}
\subsection{Wilk's Theorem}
Let the parameter $\vec{\theta}$ have $k$ dimensions.
\begin{definition}{Dimension of a hypothesis}
    The \underline{dimension} of a hypothesis $H_0 : \vec{\theta} \in \Theta_0$ is the number of free parameters in $\Theta_0$.
\end{definition}
\begin{examples}{}
    \item If $\Theta_0 = \subsetselect{\vec{\theta} \in \R^k}{\theta_1 = \theta_2 = \cdots = \theta_p = 0}$, then the dimension of the hypothesis is $k - p$.
    \item If $\Theta_0 = \subsetselect{\vec{\theta} \in \R^k}{\theta_i = f_i(\phi),~1\leq i \leq k}$, for some $\phi \in \R^{k-p}$ then the dimension of the hypothesis is again $k - p$.
\end{examples}
\begin{theorem}[Wilk's Theorem]
    Let $\Theta_0 \subseteq \Theta_1$ and $\dim(\Theta_1) - \dim(\Theta_0) = p$.

    Then given independent and identically distributed random variables $X_i$ for $i = 1$ to $n$, then under regularity conditions as $n \to \infty$ then:
    \begin{equation*}
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) \sim \chi_p^2
    \end{equation*}
    \label{thmWilks}
\end{theorem}
\begin{proof}
    This theorem will be used without proof.
\end{proof}
\begin{example}[Revisiting example~\ref{expTwoSideNormalTest}]
    In the two-sided normal mean test, we had:
    \begin{equation*}
        \Theta_0 = \{\mu_0\},\quad \Theta_1 = \R\backslash\{\mu_0\} \\
    \end{equation*}
    and we saw that:
    \begin{equation*}
        2\log{\Lambda_{\vec{X}}(H_0, H_1)} \sim \chi_1^2
    \end{equation*}
    which was an exact statement in this case. In a different parametric family, with large $n$, this would hold approximately.
\end{example}
\section{Chi-Squared Testing}
\subsection{Goodness-of-Fit Tests}
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
        Type & Count observed & Predicted freq. & Count expected \\
        \hline
        SY & $315$ & $9 / 16$ & 312 \\
        SG & $108$ & $3 / 16$ & 104 \\
        WY & $102$ & $3 / 16$ & 104 \\
        WG & $31$ & $1 / 16$ & 35 \\
        \hline
    \end{tabular}
    \caption{Table of Mendel's Data}
    \label{tabMendelExperiment}
\end{table}
\begin{example}
    In an experiment, Gregor Mendel crossed 556 smooth, yellow male peas with wrinkled, green female peas. Table~\ref{tabMendelExperiment} shows the data. In order to do a hypothesis test, let $X_1, \cdots, X_n$ be independent and identically distributed samples from a distribution on $[1, k] \cap \N$. Let $p_i = \P(X_i = i)$ and let $N_i$ be the number of observations in $\vec{X}$ of type $i$, with $\vec{N}$ having components $N_i$.

    Then $\vec{N} \sim \text{Mult}(n, p_1, \cdots, p_k)$.

    The likelihood is proportional to $p_1^{N_1} p_2^{N_2} \cdots P_k^{N_k}$, and so the log-likelihood is:
    \begin{equation*}
        l(\vec{p}) = \sum_{i=1}^{k} N_i \log(p_i) + \text{const}
    \end{equation*}
    Consider the null hypotheses to be the probabilities $\tilde{\vec{p}}$ given in the table, and the alternative to be anything else. $\Theta_1 = \subsetselect{\vec{p} \in [0, 1]^k}{\sum_i p_i = 1}$.
    \begin{equation*}
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) = 2\left(\sup_{\vec{p} \in \Theta_1} l(p) - \sup_{p \in \Theta_0} l(p)\right)
    \end{equation*}
    Note that $\sup_{\vec{p} \in \Theta_1} l(\vec{p}) = \sup_{\vec{p} \in\Theta_1} \sum_{i=1}^{k} N_i \log(p_i)$. Therefore we can use the Lagrangian,
    \begin{equation*}
        L(\vec{p}, \lambda) \sum_{i=1}^{k} N_i \log(p_i) - \lambda \left(\sum_{i=1}^{k} p_i - 1\right)
    \end{equation*}
    to show that the MLE $\hat{\vec{p}}$ is given by $\hat{p}_i = \frac{N_i}{n}$. Therefore,
    \begin{equation*}
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) = 2 \sum_{i=1}^{k} N_i \log\left(\frac{N_i}{n \tilde{p}_i}\right)
    \end{equation*}
    and now \thmref{thmWilks} says that $2\log(\Lambda) \sim \chi_p^2$. note that our dimension for $\Theta_1$ is $k - 1$ (because we have $k$ dimensions and 1 linear condition), and the dimension for $\Theta_0$ is $0$. Therefore we reject $H_0$ if $2\log(\Lambda) \geq \chi_{k-1}^2(\alpha)$. It is common to write:
    \begin{equation*}
        2\log(\Lambda) = \sum_{i=1}^{n} o_i \log\left(\frac{o_i}{e_i}\right)
    \end{equation*}
    where $o_i$ is the observed number of type $i$, $N_i$, and $e_i$ is the expected number of type $i$, $n \tilde{p}_i$.
    
    Then we let $\delta_i = o_i - e_i$. Then:
    \begin{align*}
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) &= 2\sum_{i=1}^{n} (e_i + \delta_i) \log\left(1 + \frac{\delta_i}{e_i}\right) \\
        &= 2 \sum_{i=1}^{n} (e_i + \delta_i) \left(\frac{\delta_i}{e_i} - \frac{\delta_i^2}{2e_i^2} + \cdots\right) \\
        &\approx \sum_{i=1}^{n} \frac{\delta_i^2}{e_i} \\
        &= \sum_{i=1}^{n} \frac{(o_i - e_i)^2}{e_i}
    \end{align*}
    which is Pearson's Chi-Squared Statistic.

    In the pea plant experiment,
    \begin{equation}
        \sum_{i=1}^{4} \frac{(o_i - e_i)^2}{e_i} \approx 0.604
    \end{equation}
    and $\chi_3^2(0.05) \approx 7.815 > 0.604$. Therefore we do not reject $H_0$. The $p$-value is:
    \begin{equation*}
        \P(X_3^2 \geq 0.604) \approx 0.96
    \end{equation*}
\end{example}
\subsection{Composite Null Hypotheses}
Suppose we wish to test $H_0 : p_i = p_i(\theta)$ for some $\theta$ against $H_1 : \vec{p} \in \subsetselect{\vec{p} \in [0, 1]^k}{\sum_{i=1}^{k} p_i = 1}$.
\begin{example}
    Consider $H_0$ to be:
    \begin{equation*}
        \vec{p} = \begin{pmatrix}\theta^2 \\ 2\theta(1-\theta) \\ (1-\theta)^2\end{pmatrix}, \theta \in [0, 1]
    \end{equation*}
    and then the GLRT is:
    \begin{align*}
        2\log(\Lambda_{\vec{X}}(H_0, H_1)) &= 2 \left(\sup_{\sum_i p_i = 1} l(\vec{p}) - \sup_\theta l(\vec{p}(\theta))\right) \\
        &= 2 \left(l(\hat{p}) - l(\vec{p}(\hat{\theta}))\right) \\
        &= 2\sum_{i=1}^{3} N_i \log\left(\frac{N_i}{n p_i(\hat{\theta})}\right)
        &= \sum_{i=1}^{3} o_i \log\left(\frac{o_i}{e_i}\right) \\
        &\approx \frac{(o_i - e_i)^2}{e_i}
    \end{align*}
    and so so one can perform a test with $\chi_p^2$ where $p = \dim(\Theta_1) - \dim(\Theta_0) = 2 - 1 = 1$.
\end{example}
\section{Contingency Tables}
\subsection{Testing for Independence}
Suppose $(X_1, Y_1), \cdots, (X_n, Y_n)$ are independent and identically distributed with $X$ taking values in $\{1, \cdots, r\}$ and $Y_1$ taking values in $\{1, \cdots, c\}$. We wish to test the null hypothesis that $X_i$ and $Y_i$ are independent. The entries in a \underline{contingency table} are:
\begin{equation*}
    N_{ij} = \left|\subsetselect{l}{1 \leq l \leq n, (X_l, Y_l) = (i, j)}\right|
\end{equation*}
That is, $N_{ij}$ is the number of samples of type $(i, j)$.
\begin{table}
    \centering
    \begin{tabular}{|c|c|c|c|}
        \hline
         & F & M & Other \\
        \hline
        Applied & 10 & 9 & 1 \\
        TP & 11 & 59 & 0 \\
        Pure & 10 & 56 & 2 \\
        Stats & 6 & 23 & 0 \\
        Part II & 13 & 87 & 1 \\
        \hline
    \end{tabular}
    \caption{Table of admissions data to part III}
    \label{tabAdmissions}
\end{table}
\begin{example}
    Consider table~\ref{tabAdmissions}. We wish to test the null hypothesis that gender and stream are independent.

    We observe $n$ samples. A sample of type $(i, j)$ has probability $p_{ij}$. Then:
    \begin{equation*}
        (N_{11}, N_{12}, \cdots, N_{1c}, N_{21}, \cdots, N_{rc}) \sim \text{Mult}(n, p_{11}, p_{12}, \cdots, p_{rc})
    \end{equation*}
    Define:
    \begin{equation*}
        p_{i+} = \sum_{j= 1}^{c}p_{ij},\quad p_{+j} = \sum_{i = 1}^{r} p_{ij}
    \end{equation*}
    then the hypotheses are:
    \begin{align*}
        H_0: & p_{ij} = p_{i+} \times p_{+j}~\forall i, j \\
        H_1: & \{p_{ij}\} \text{ constrained only by } \sum_{i, j}^{p_{ij}} = 1
    \end{align*}
    Then using Lagrangian methods, the MLEs are $\hat{p}_{ij} = \frac{N_{ij}}{n}$ under $H_1$ and $(\hat{p}_{i+}, \hat{p}_{+j}) = \left(\frac{N_{i+}}{n}, \frac{N_{+j}}{n}\right)$.

    Therefore, the GLR statistic is:
    \begin{equation*}
        2\log(\Lambda) = 2\sum_{i=1}^{r} \sum_{j=1}^{c} N_{ij} \log\left(\frac{\hat{p}_{ij}}{\hat{p_{i+} \hat{p}_{+j}}}\right)
    \end{equation*}
    writing $o_{ij} = N_{ij}$ and $e_{ij} = \hat{p}_{i+} \hat{p}_{+j}$,
    \begin{align*}
        2\log(\Lambda) &= 2\sum_{i=1}^{r} \sum_{j=1}^{c} o_{ij} \log\left(\frac{o_{ij}}{e_{ij}}\right) \\
        &\approx 2\sum_{i=1}^{r} \sum_{j=1}^{c} \frac{(o_{ij} - e_{ij})^2}{e_{ij}}
    \end{align*}
    Then by \thmref{thmWilks}, this converges to a $\chi^2$ distribution. The number of degrees of freedom is:
    \begin{equation*}
        \dim(\Theta_1) - \dim(\Theta_0) = (rc - 1) - ((r - 1) + (c - 1)) = (r-1)(c-1)
    \end{equation*}

    Then we find:
    \begin{equation*}
        \sum_{i=1}^{3} \sum_{j=1}^{5} \frac{(o_{ij} - e_{ij})^2}{e_{ij}} \approx 22.2
    \end{equation*}
    and the $p$-value is:
    \begin{equation*}
        \P(\chi_8^2 \geq 22.2) \approx 0.005
    \end{equation*}
    which is significant, and so we reject the hypothesis that these are independent.
\end{example}
\subsection{Testing for Homogeneity}
\begin{example}
    Suppose 150 patients are randomly allocated to three groups of equal size. One group is given a placebo, one a half-dose of a drug, and the last a full dose of the drug.

    We cannot use the methods in the last subsection because here the row totals are fixed. We want to test the null hypothesis of homogeneity, that is, $H_0$ corresponds to the probability of response being unchanged between groups. That is, the drug has no effect.

    Suppose that row $i$ has a fixed number of entries $n_{i+}$. Then for probabilities $p_{ij}$,
    \begin{equation*}
        N_{i1}, \cdots, N_{ic} \sim \text{Mult}(n_{i+}, p_{i1}, \cdots, p_{ic})
    \end{equation*}
    \begin{align*}
        H_0:& p_{1j} = p_{2j} = \cdots, = p_{rj}~\forall j \in \{1, \cdots, c\} \\
        H_1:& \sum_{i=1}^{r} \sum_{j=1}^{c} p_{ij} = 1
    \end{align*}

    Then we use a GLRT. Under $H_1$, our likelihood is:
    \begin{equation*}
        L(p) = \prod_{i=1}^{r} \frac{n_{i+}!}{N_{i1}! \cdots, N_{ic}!} p_{i1}^{N_{i1}} \cdots p_{ic}^{N_{ic}}
    \end{equation*}
    and the log-likelihood is:
    \begin{equation*}
        l(p) = \text{const} + \sum_{i=1}^{r} \sum_{j=1}^{c} N_{ij} \log(p_{ij})
    \end{equation*}
    Using the Lagrangian method with the constraint that the $p_{ij}$ sum to $1$, the MLE is:
    \begin{equation*}
        \hat{p}_{ij} = \frac{N_{ij}}{n_{i+}}
    \end{equation*}

    Under $H_0$, let $p_{ij} = p_i$. Then
    \begin{equation*}
        l(p) = \text{const} + \sum_{i=1}^{r} \sum_{j=1}^{c} N_{ij} \log(p_j) = \sum_{j=1}^{c} N_{+j} \log(p_j)
    \end{equation*}
    Now using the Lagrangian method with the constraint $\sum_{j=1}^{c} p_j = 1$, the MLE is
    \begin{equation*}
        \hat{p}_j = \frac{N_{+j}}{n_{+j}}
    \end{equation*}
    and so our GLR statistic is:
    \begin{align*}
        2\log(\Lambda) &= 2\sum_{i=1}^{r}\sum_{j=1}^{c} N_{ij} \log\left(\frac{\hat{p}_{ij}}{\hat{p}_j}\right) \\
        &= 2\sum_{i=1}^{r}\sum_{j=1}^{c} N_{ij} \log\left(\frac{N_{ij}}{n_{i+} N_{+j} / n_{+j}}\right) \\
        &= 2\sum_{i=1}^{r}\sum_{j=1}^{c} N_{ij} \frac{(o_{ij} - e_{ij})^2}{e_{ij}}
    \end{align*}
    where $o_{ij} = N_{ij}$ and $e_{ij} = \frac{n_{i+ N_{+j}}}{n_{+j}}$.

    Our degrees of freedom is:
    \begin{equation*}
        \dim(\Theta_1) - \dim(\Theta_0) = r(c-1) - (c-1) = (r-1)(c-1)
    \end{equation*}
    and so our asymptotic $\chi_p^2$ distribution is the same as the independence test.
\end{example}
\section{Facts about Normal Random Variables}
\subsection{Multivariate Normal Theory}
Let $\vec{X} = (X_1, \cdots, X_n)^T$ be a vector of random variables. Recall that the expectation is a vector and the variance/covariance is a matrix.
\begin{align*}
    \left[\E[\vec{X}]\right]_i &= \E[X_i] \\
    \left[\Var(\vec{X})\right]_{ij} &= \Cov(X_i, X_j)
\end{align*}

For a matrix $A \in \R^{m \times n}$ and $\vec{b} \in \R^m$,
\begin{equation*}
    E[A\vec{X} + \vec{b}] = AE[\vec{X}] + \vec{b},\quad \Cov(A\vec{X} + \vec{b}) = A\Cov(X)A^T
\end{equation*}
\begin{definition}{Multivariate normal distribution}
    A random vector $\vec{X}$ has a \underline{multivariate normal distribution} if, for any $\vec{t} \in \R^n$, the random variable $\vec{t} \cdot \vec{X}$ has a normal distribution.
\end{definition}
\begin{proposition}
    If $\vec{X}$ is a multivarite normal vector, then $A\vec{X} + \vec{b}$ is a multivariate normal for any $A \in \R^{m \times n}$ and $\vec{b} \in \R^m$.
    \label{propTransformMVN}
\end{proposition}
\begin{proof}
    Take any $\vec{t} \in \R^m$. Then:
    \begin{align*}
        \vec{t} \cdot \left(A\vec{X} + \vec{b}\right) &= (A^T \vec{t})^T \vec{X} + \vec{t} \cdot \vec{b} \\
    \end{align*}
    Then since $\vec{X}$ is a multivariate normal, $(A\vec{t}) \cdot \vec{X}$ has a normal distribuition. Adding a constant ($\vec{t} \cdot \vec{b}$) shifts the distribution but it is still normal.
\end{proof}
\begin{proposition}
    A multivariate normal distribution is fully specified by its mean $\vec{\mu}$ and covariance matrix.
    \label{propMVNParameters}
\end{proposition}
\begin{proof}
    Let $\vec{X_1}$ and $\vec{X_2}$ be multivariate normal. Let them both have mean $\vec{\mu} \in \R^n$ and covariance matrix $\Sigma \in \R^{n \times n}$. The moment generating functions are:
    \begin{align*}
        M_{\vec{X_1}}(t) &= \E\left[e^{\vec{t} \cdot \vec{X_1}}\right] \\
        &= \exp\left(\vec{t} \cdot \vec{\mu} + \frac12 \vec{t} \cdot \Sigma \vec{t}\right) \text{ by recognising normal dist.}
    \end{align*}
    and this is the same for $\vec{X_2}$, so the random variables must have the same distribution.
\end{proof}
\begin{corollary}
    Suppose $X_1, \cdots, X_n$ are independent random variables with $X_i \sim N(\mu_i, \sigma^2)$. Let $\vec{X}$ be the vector of these random variables, and $\vec{\mu}$ be the vector of the means.

    Then for an orthogonal matrix $A$, the elements of $\vec{Y} = A\vec{X}$ are independently distributed, with $Y_i \sim N((A\vec{\mu})_i, \sigma^2)$
    \label{corRotateNormals}
\end{corollary}
\begin{proof}
    The distribution of $\vec{Y}$ is determined by its mean and covariance.

    \begin{equation*}
        \E[\vec{Y}] = \E[A\vec{X}] = A\E[\vec{X}] = A\vec{\mu}.
    \end{equation*}

    \begin{align*}
        \Cov(A\vec{X}) &= A\Cov(\vec{X})A^T \\
        &= A(\sigma^2 I)A^T \\
        &= \sigma^2 AA^T \\
        &= \sigma^2 I
    \end{align*}
\end{proof}
\subsection{Distribution of \texorpdfstring{$\overline{X}$}{X-bar} and \texorpdfstring{$S_{xx}$}{Sxx}}
Suppose we have a normal distribution with unknown mean and variance. Then the samples $X_i$ are independent and identically distributed, $X_i \sim N(\mu, \sigma^2)$. Then the MLE for $\mu$ is:
\begin{equation*}
    \hat{\mu} = \overline{X} = \frac1n \sum_{i=1}^{n} X_i
\end{equation*}
and the MLE for $\sigma^2$ is:
\begin{equation*}
    \hat{\sigma}^2 = \frac{S_{xx}}{n} = \frac1n \sum_{i=1}^{n} (X_i - \overline{X})^2
\end{equation*}
\begin{theorem}
    With the above setup,
    \begin{enumerate}
        \item $\overline{X} \sim N\left(\mu, \frac{\sigma^2}{n}\right)$
        \item $\frac{S_{xx}}{\sigma^2} \sim \chi_{n-1}^2$
        \item $\overline{X}$ and $S_{xx}$ are independent.
    \end{enumerate}
    \label{thmNormalStatistics}
\end{theorem}
\begin{proof}
    Part 1 is an easy consequence of the fact that linear combinations of Gaussians are Gaussian.

    For parts 2 and 3, write:
    \begin{align*}
        \sum_{i=1}^{n} (X_i - \mu)^2 &= \sum_{i=1}^{n} \left((X_1 - \overline{X}) + (\overline{X} - \mu)\right)^2 \\
        &= \sum_{i=1}^{n} \left((X_i - \overline{X})^2 + 2(X_i - \overline{X})(\overline{X} - \mu) + (\overline{X} - \mu)^2\right) \\
        &= S_{xx} + n(\overline{X} - \mu)^2
    \end{align*}
    Let $A$ be an orthogonal matrix such that:
    \begin{align*}
        \vec{Y} &= A\left(\vec{X} - \begin{pmatrix}\mu \\ \vdots \\ \mu\end{pmatrix}\right) = \begin{pmatrix}\sqrt{n}(\overline{X} - \mu) \\ Y_2 \\ \vdots \\ Y_n\end{pmatrix} \\
        \therefore A &=
        \begin{pmatrix}
            1/\sqrt{n} & 1/\sqrt{n} & \cdots & 1/\sqrt{n} \\
             & & & \\
             & \large\tilde{A} & &
        \end{pmatrix}
    \end{align*}
    where $\tilde{A}$ is chosen such that $A$ is orthogonal.

    Then $Y_1 = \sqrt{n}(\overline{X} - \mu) \sim N(0, \sigma^2)$ and is independent of the rest of the $Y_i$ by \thmref{corRotateNormals}.
    \begin{align*}
        \norm{\vec{Y}}^2 &= \norm{A(\vec{X} - \vec{\mu})}^2 \\
        &= (\vec{X} - \vec{\mu})^T A^T A (\vec{X} - \vec{\mu}) \\
        &= (\vec{X} - \vec{\mu})^T (\vec{X} - \vec{\mu}) \\
    \end{align*}
    which means,
    \begin{equation*}
        \sum_{i=1}^{n} Y_i^2 = \sum_{i=1}^{n} (X_i - \mu)^2
    \end{equation*}
    and so we must have:
    \begin{equation*}
        \sum_{i=2}^{n} Y_i^2 = \sum_{i=1}^{n} (X_i - \mu)^2 - n(\overline{X} - \mu)^2 = S_{xx}
    \end{equation*}
    so $S_{xx}$ is independent of $Y_1 = \overline{X}$.

    Further,
    \begin{equation*}
        \frac{S_{xx}}{\sigma^2} = \sum_{i=2}^{n} \frac{Y_1^2}{\sigma^2} \sim \chi_{n-1}^2
    \end{equation*}
\end{proof}
\section{\texorpdfstring{$t$}{T}-tests}
\begin{definition}{$t$-distribution}
    If $U \sim N(0, 1)$ and $V \sim \chi_n^2$, with $U$ and $V$ independent, then
    \begin{equation*}
        T = \frac{U}{\sqrt{V / n}}
    \end{equation*} has a \underline{$t$-distribution} with $n$ degrees of freedom. We write:
    \begin{equation*}
        T \sim t_n.
    \end{equation*}
\end{definition}
\begin{example}
    Suppose that $X_i$ are independent and identically distributed with $X_i \sim N(\mu, \sigma^2)$. When $\sigma^2$ is known, we find a $100(1-\alpha)\%$ confidence interval for $\mu$ by noting that:
    \begin{equation*}
        Z = \frac{\sqrt{n}(\overline{X} - \mu)}{\sigma} \sim N(0, 1).
    \end{equation*}
    If instead $\sigma^2$ is unknown, we note:
    \begin{equation*}
        \overline{X} \sim N(\mu, \sigma^2 / n),\quad \frac{S_{xx}}{\sigma^2} \sim \chi_{n-1}^2
    \end{equation*}
    and then find a $t$-distribution:
    \begin{equation*}
        T = \frac{\sqrt{n}(\overline{X} - \mu) / \sigma}{\sqrt{S_{xx} / \sigma^2(n-1)}} \sim t_{n-1}
    \end{equation*}
    and so we can find a confidence interval as follows:
    \begin{equation*}
        \P_{\mu, \sigma^2} \left(-t_{n-1}\left(\frac{\alpha}{2}\right) \leq \frac{\sqrt{n} (\overline{X} - \mu)}{\sqrt{S_{xx} / (n-1)}} \leq t_{n-1} \left(\frac{\alpha}{2}\right)\right) = 1- \alpha
    \end{equation*}
    and so,
    \begin{equation*}
        \overline{X} \pm \sqrt{\frac{S_{xx}}{n-1}} \frac{t_{n-1}\left(\frac{\alpha}{2}\right)}{\sqrt{n}}
    \end{equation*}
    is an exact $100(1-\alpha)\%$ confidence interval for $\mu$.

    If we instead wish to test $H_0 : \mu = \mu_0$ with $\sigma^2$ unknown, then under $H_0$,
    \begin{equation*}
        T = \frac{\sqrt{n}(\overline{X} - \mu_0)}{\sqrt{S_{xx} / (n-1)}} \sim t_{n-1}
    \end{equation*}
    so reject $H_0$ when $\abs{T} \geq t_{n-1}\left(\frac{\alpha}{2}\right)$. This is a \underline{$t$-test}.
\end{example}
\begin{remarks}
    \item As $n \to \infty$, $\frac{S_{xx}}{n-1} \to \sigma^2$ and $t_{n-1}\left(\frac{\alpha}{2}\right) \to z_{\alpha / 2}$. Then the $t$-test converges to a $z$-test. Importantly, the $t$-test does not need large $n$ to be valid.
    \item The $t$-test can be seen as an example of a GLRT for:
        \begin{align*}
            H_0:& \mu = \mu_0 \\
            H_1:& \mu \neq \mu_0
        \end{align*}
        The GLR statistic is:
        \begin{equation*}
            \Lambda_{\vec{X}}(H_0, H_1) = \frac{\sup_{\mu, \sigma^2} f(\vec{x}, \mu, \sigma^2)}{\sup_{\sigma^2} f(\vec{x}, \mu_0, \sigma^2)}
        \end{equation*}
        and substituting $(\hat{\mu}, \hat{\sigma}^2) = (\overline{X}, S_{xx} / n)$ gives the numerator:
        \begin{equation*}
            \left(\frac{2\pi\sum_{i=1}^{n} (X_i - \overline{X})^2}{n}\right)^{-n / 2} \exp\left(-\frac{n}{2}\right)
        \end{equation*}
        and when $\mu = \mu_0$, the MLE for $\sigma^2$ is $\frac1n \sum_{i=1}^{n} (X_i - \mu_0)^2$. The denominator is then:
        \begin{equation*}
            \left(\frac{2\pi\sum_{i=1}^{n} (X_i - \mu_0)^2}{n}\right)^{-n / 2} \exp\left(-\frac{n}{2}\right)
        \end{equation*}
        and so the GLR statitsic is:
        \begin{align*}
            \Lambda_{\vec{X}}(H_0, H_1) &= \frac{\left(\frac{2\pi \sum_{i=1}^{n} (X_i - \overline{X})^2}{n}\right)^{-n / 2}}{\left(\frac{2\pi \sum_{i=1}^{n} (X_i - \mu_0)^2}{n}\right)^{-n / 2}} \\
            &= \left(1 + \frac{n(\overline{X} - \mu_0)^2}{\sum_{i=1}^{n} (X_i - \overline{X})^2}\right)^{-n/2}
        \end{align*}
        and so we reject $H_0$ when
        \begin{equation*}
            \frac{\sqrt{n} (\overline{X} - \mu_0)}{\sqrt{S_{xx} / (n-1)}}
        \end{equation*}
        is large. This gives another idea of where the $t$-distribution comes from.
\end{remarks}
\section{Analysis of Variance (ANOVA)}
We consider testing equality of means among $k > 2$ groups. The model is:
\begin{equation*}
    X_{ij} = \mu_i + \epsilon_{ij},\quad 1\leq j \leq n_i, 1 \leq i \leq k
\end{equation*}
where $n_i$ is the number observations for the $j$th group.

Let $N = \sum_{i=1}^{k} n_i$ and assume $\epsilon_{ij} \sim N(0, \sigma^2)$ are independent. We test:
\begin{align*}
    H_0: & \mu_1 = \mu_2 = \cdots = \mu_k \\
    H_1: & \text{ not $H_1$}
\end{align*}
We use a generalised likelihood ratio test, with likelihood ratio statistic:
\begin{align*}
    \Lambda_{\vec{X}}(H_0, H_1) &= \frac{\sup_{\mu_1, \cdots, \mu_k, \sigma^2} (2\pi\sigma^2)^{-N / 2} \exp\left[-\sum_{i=1}^{k} \sum_{j=1}^{n_i} \frac{(X_{ij} - \mu_i)^2}{2\sigma^2}\right]}{\sup_{\mu, \sigma^2} (2\pi\sigma^2)^{-N / 2} \exp\left[-\sum_{i=1}^{k} \sum_{j=1}^{n_i} \frac{(X_{ij} - \mu)^2}{2\sigma^2}\right]} \\
    &= \left(\frac{s_0}{s_1}\right)^{N/2}
\end{align*}
where
\begin{align*}
    s_0 &= \sum_{i=1}^k \sum_{j=1}^{n_i} (X_{ij} - \overline{X}_{\cdot\cdot})^2 \\
    s_1 &= \sum_{i=1}^k \sum_{j=1}^{n_i} (X_{ij} - \overline{X}_{i\cdot})^2 \\
\intertext{and}
    \overline{X}_{\cdot\cdot} &= \frac{\sum_{i=1}^k \sum_{j=1}^{n_i} X_{ij}}{N} \\
    \overline{X}_{i\cdot} &= \frac{\sum_{j=1}^{n_i} X_{ij}}{N_i}
\end{align*}
so we should reject $H_0$ if $\frac{s_0}{s_1}$ is large, and note that:
\begin{align*}
    s_0 &= \sum_{i=1}^k \sum_{j=1}^{n_i} (X_{ij} - \overline{X}_{i\cdot} + \overline{X}_{i\cdot} - \overline{X}_{\cdot\cdot})^2 \\
    &= \sum_{i=1}^k \sum_{j=1}^{n_i} \left(\left(X_{ij} - \overline{X}_{i\cdot}\right)^2 + 2\left(X_{ij} - \overline{X}_{i\cdot}\right)\left(\overline{X}_{i\cdot} - \overline{X}_{\cdot\cdot}\right) + \left(\overline{X}_{i\cdot} - \overline{X}_{\cdot\cdot}\right)\right) \\
    &= \underbrace{\sum_{i=1}^k \sum_{j=1}^{n_i} \left(X_{ij} - \overline{X}_{i\cdot}\right)^2}_{s_1} + \underbrace{\sum_{i=1}^{k} n_i \left(\overline{X}_{i\cdot} - \overline{X}_{\cdot\cdot}\right)}_{s_2}
\end{align*}
and we conclude that rejecting when $\frac{s_0}{s_1}$ is large is equivalent to rejecting when $\frac{s_2}{s_1}$ is large. $s_1$ is the ``within samples sum of squares'', and $s_2$ the ``between samples sum of squares''. Next, we derive the distribution of this. Note that,
\begin{equation*}
    \sum_{j=1}^{n_i} \left(X_{ij} - \overline{X}_{i\cdot}\right)^2 \text{ is independent of $\overline{X}_{i\cdot}$}
\end{equation*}
and this gives us that $s_1$ is independent of $s_2$. We can see this by considering $s_2$ as a function of $\{\overline{X}_{i\cdot}\}_i$ and the observations are independent for different $i$.

Also, under $H_0$, $S_1 \sim \sigma^2 \chi_{N-k}^2$ and $S_2 \sim \sigma^2 \chi_{k-1}^2$. We find the degrees of freedom for $S_1$ because:
\begin{equation*}
    \sum_{i=1}^{k} (n_i - 1) = N-k
\end{equation*}
and for $S_2$ we have $k-1$ because it is the sample variance of $k$ independent and identically distributed $N(0, \sigma^2)$ random variables. We now need to know what the quotient of chi-squared distributions is.
\begin{definition}{$F$-distribution}
    If $V \sim \chi_n^2$ and $W\sim\chi_m^2$, with $V$ and $W$ independent, we say that $F = \frac{V/n}{W/m}$ has an \underline{$F$-distribution} with $(n, m)$ degrees of freedom. This is also denoted $F_{n, m}$.
\end{definition}
Using this definition, we see that, under $H_0$,
\begin{equation*}
    \frac{S_2 / (k-1)}{S_1 / (N-k)} \sim F_{k-1, N-1}
\end{equation*}
and we can construct a test of size $\alpha$.

\begin{remark}
    Recall the $t$ distribution corresponding to:
    \begin{equation*}
        T = \frac{U}{\sqrt{V / n}},\quad U \sim N(0, 1), V\sim\chi_n^2
    \end{equation*}
    so since $U^2 \sim \chi_1^2$, the statistic:
    \begin{equation*}
        T^2 = \frac{U^2 / 1}{V / n} \sim F_{1,n}
    \end{equation*}
\end{remark}
\begin{table}
    \begin{tabular}{|c|c|c|c|c|c|}
        \hline
        College & Students & Total marks & Probability & DEs & VC \\
        \hline
        1 & 44 & 318 & 36 & 29 & 32 \\
        2 & 18 & 283 & 25 & 37 & 35 \\
        3 & 16 & 294 & 28 & 35 & 37 \\
        4 & 15 & 244 & 15 & 39 & 31 \\
        5 &  7 & 280 & 24 & 34 & 35 \\
        \hline
    \end{tabular}
    \caption{Table of 2024-25 IA results}
    \label{tabANOVA}
\end{table}
\begin{example}
    Consider table~\ref{tabANOVA}, and we would like to perform analysis of variance on this. We have $k = 5$, $N = 100$, so our $F$ distribution is $F_{4,95}$.

    At a $5\%$ significance level, we reject $H_0$ if the statistic is larger than:
    \begin{equation*}
        F_{4, 95}(0.05) \approx 2.47
    \end{equation*}
    For the total marks, the $F$ statistic is $3.71$ so reject $H_0$. For IA Probability, the $F$ statistic is $4.25$ so reject $H_0$. For IA Differential Equations, the $F$ statistic is $1.70$ so do not reject $H_0$. For IA Vector Calculus, the $F$ statistic is $0.50$ so do not reject $H_0$.
\end{example}
\begin{remark}
    What we described is sometimes called 1-way ANOVA. In practice, 2-way ANOVA analyses data which are defined according to 2 variables. For example, we could have considered students' exam performance based on whether supervisions were given and monetary incentive offered. Here we would consider:
    \begin{equation*}
        Y_{ijk} = \alpha + \mu_j + \lambda_k + \epsilon_{ijk}
    \end{equation*}
    where $(j, k)$ is a binary coding of supervision/money.

    The test is then for:
    \begin{align*}
        H_0: & \mu_1 = \lambda_1 = 0 \\
        H_1: & \mu_1, \lambda_1 \in \R
    \end{align*}
\end{remark}
\end{document}