\documentclass[../Main.tex]{subfiles}

\begin{document}
\section{Faraday's Law of Induction}
We are interested in the third Maxwell equation:
\begin{equation*}
    \nabla \times \vec{E} = -\frac{\partial B}{\partial t} \tag{\ref{eqnMEC}}.
\end{equation*}
This implies that a time-dependent magnetic field must be accompanied by an electric field. This can induce a current to flow in a conductor, a process known as electromagnetic induction.
\subsection{Faraday's Law for a Static Circuit}
Consider a closed curve $C$ that is the boundary of a time-independent open surface $C$. We integrate \eqnref{eqnMEC} over $S$ and use Stokes's Theorem:
\begin{equation*}
    \int_C \vec{E} \cdot \vec{dx} = -\int_S \frac{\partial \vec{B}}{\partial t} \cdot \vec{dS} = - \frac{d}{dt} \int_S \vec{B} \cdot \vec{dS}
\end{equation*}
This is \underline{Faraday's law of Induction}. We also write it:
\begin{equation*}
    \emf = -\frac{d\flux}{dt}
\end{equation*}
Here
\begin{equation*}
    \emf = \int_C \vec{E} \cdot \vec{dx}
\end{equation*}
is the \underline{electromotive force} (emf) around $C$ and
\begin{equation*}
    \flux = \int_S \vec{B} \cdot \vec{dS}
\end{equation*}
is the \underline{magnetic flux} through $S$.

Since $\nabla \cdot \vec{B} = 0$, the flux $\flux$ is the same for any surface $S$ such that $\partial S = C$, and so $\flux$ can be understood to be a property of $C$. In this case, we say $\flux$ is the flux through $C$.

Using $\vec{B} = \nabla \times \vec{A}$ and Stokes's Theorem, we can write
\begin{equation*}
    \flux = \int_C \vec{A} \cdot \vec{dx}
\end{equation*}
which is invariant under a gauge transformation $\tilde{\vec{A}} = \vec{A} + \nabla \chi$. This makes the fact that $\flux$ depends on $C$ and not $S$ clear.

The emf is not actually a force. It is the line integral of the Lorentz force on a particle of unit charge confined to $C$:
\begin{equation*}
    \emf = \frac1q \int_C \vec{F} \cdot \vec{dx} = \int_C (\vec{E} + \dvec{x} \times \vec{B}) \cdot \vec{dx} = \int_C \vec{E} \cdot \vec{dx}
\end{equation*}
since $\dvec{x}$ is tangent to $C$ for a particle confined to $C$ (and assuming $C$ is time-independent).

We will see later that if $C$ coincides with a thin wire of resistance $R$, then the current induced in the wire is $I = \emf / R$.

There are several ways in which the magnetic flux through $C$ could change in time:
\begin{itemize}
    \item a magnet is moved near $C$;
    \item a current-carrying circuit is moved near $C$;
    \item the current flowing in a nearby circuit is changed.
\end{itemize}
All these will induce an emf around $C$ and cause a current to flow.
\subsection{Faraday's Law for a Moving Circuit}
Now let $C(t)$ be a \underline{time-dependent} closed curve that is the boundary of an open surface $C(t)$. Then we want to understand how the magnetic flux,
\begin{equation*}
    \flux(t) = \int_{S(t)} \vec{B}(\vec{x}, t) \cdot \vec{dS}
\end{equation*}
changes in time.

Consider a small time interval $\delta t$ and expand the difference in $\flux$ to first order.
\begin{align}
    \flux(t + \delta t) &- \flux(t) = \int_{S(t + \delta t)} \vec{B}(\vec{x}, t + \delta t) \cdot \vec{dS} - \int_S \vec{B}(\vec{x}, t) \cdot \vec{dS} \nonumber \\
    &= \int_{S(t + \delta t)} \left(\vec{B}(\vec{x}, t) + \frac{\partial \vec{B}}{\partial t} \delta t + O(\delta t^2)\right) \cdot \vec{dS}\nonumber\\
    &\qquad- \int_{S(t)} \vec{B}(\vec{x}, t) \cdot \vec{dS} \label{eqnFluxDifference} \\
    &= \int_{S(t + \delta t) \backslash S(t)} \vec{B}(\vec{x}, t) \cdot \vec{dS} + \int_{S(t)} \frac{\partial \vec{B}}{\partial t} \cdot \vec{dS}~\delta t + O(\delta t^2) \nonumber
\end{align}
To resolve the first term, let $\delta V$ be the volume swept out by $S(t)$ in the time interval $\delta t$. Its boundary is the closed surface $S(t + \delta t) \backslash S(t) \cup \Sigma$ where $\Sigma$ is the surface swept out by $C(t)$ in the $\delta t$. See figure~\ref{figTimeDepSurface}.

\begin{figure}
    \centering
    \begin{tikzpicture}
        \tikzset{linearrow/.pic={
        \draw (-0.1, -0.1) -- (0, 0) -- (0.1, -0.1);
        }}
        \draw[blue] (0, 0) ellipse[x radius=5, y radius=2];
        \draw[blue] (0, 5) ellipse[x radius=5, y radius=2];
        \draw[blue] (-5, 0) -- (-5, 5);
        \draw[blue] (5, 0) -- (5, 5);
        
        \draw (0, -2) pic[rotate=-90] {linearrow}
            node[anchor=north] {$C(t)$};
        \draw (0, 3) pic[rotate=-90] {linearrow}
            node[anchor=north] {$C(t + \delta t)$};

        \node at (0, 0) {$S(t)$};
        \node at (0, 5) {$S(t + \delta t)$};

        \foreach \y in {0, 5} {
            \draw[->] (-2, \y) -- ++(0, 1) node[anchor=south] {$\vec{n}$};
            \draw (-1.8, \y) -- ++(0, 0.2) -- ++(-0.2, 0);
        }

        \node[anchor=west] at (5, 2.5) {$\Sigma$};

        \pgfmathsetmacro{\arcx}{5 * cos(-40)}
        \pgfmathsetmacro{\arcxcentre}{5 * cos(-35)}
        \pgfmathsetmacro{\arcy}{2 * sin(-40)}
        \pgfmathsetmacro{\toparcy}{\arcy + 5}

        \draw[thick] (\arcx, \arcy) arc[start angle=-40, end angle=-30, x radius=5, y radius=2]
            pic[pos=0.5, rotate=-70] {linearrow}
            node[pos=0.5, anchor=north] {$\vec{dx}$};

        \draw[thick] (\arcx, \arcy) -- (\arcx, \toparcy)
        pic[pos=0.5] {linearrow}
        node[pos=0.5, anchor=east] {$\vec{v} dt$}
        arc[start angle=-40, end angle=-30, x radius=5, y radius=2] -- ++(0, -5);

        \draw[->] (\arcxcentre, 1) -- ++(-30:2)
            node[anchor=west] {$\vec{dS}$};
        
    \end{tikzpicture}
    \caption{Diagram of a time-dependent surface}
    \label{figTimeDepSurface}
\end{figure}

Then by \eqnref{eqnMBD},
\begin{align*}
    0 &= \int_{\delta V} \nabla \cdot \vec{B} dV \\
    &= \int_{S(t + \delta t) \backslash S(t)} \vec{B} \cdot \vec{dS} + \int_\Sigma \vec{B} \cdot \vec{dS} \\
\end{align*}
to evaluate the last term, parameterise $C$ as $\vec{x} = \vec{x}(\lambda, t)$ where $\lambda$ is a parameter around $C$. An element of $C$ is:
\begin{equation*}
    \vec{dx} = \frac{\partial \vec{x}}{\partial \lambda} d\lambda
\end{equation*}
and has velocity $\vec{v} = \frac{\partial \vec{x}}{\partial t}$.

In time $\delta t$ it sweeps out the vector area element:
\begin{equation*}
    \vec{dS} = \vec{dx} \times (\vec{v} \delta t)
\end{equation*}
and by considering the diagram, this points out of $\delta V$, are required. Therefore:
\begin{equation*}
    \int_\Sigma \vec{B} \cdot \vec{dS} = \int_C \vec{B} \cdot \left(\vec{dx} \times \vec{v}\right) \delta t + O(\delta t^2).
\end{equation*}
Returning to \eqnref{eqnFluxDifference}:
\begin{align*}
    \flux(t + \delta t) - \flux(t) &= - \int_C (\vec{v} \times \vec{B}) \cdot \vec{dx}~\delta t + \int_S \frac{\partial \vec{B}}{\partial t} \cdot \vec{dS}~\delta t + O(\delta t^2) \\
    \frac{d\flux}{dt} &= -\int_C (\vec{v} \times \vec{B}) \cdot \vec{dx} + \int_S \frac{\partial \vec{B}}{\partial t} \cdot \vec{dS} \\
    &= -\int_C (\vec{v} \times \vec{B}) \cdot \vec{dx} - \int_S (\nabla \times \vec{E}) \cdot \vec{dS} \\
    &= -\int_C (\vec{E} + \vec{v} \times \vec{B}) \cdot \vec{dx}
\end{align*}
Therefore we recover Faraday's Law, $\emf = -\frac{d\flux}{dt}$ with the redefined emf:
\begin{equation*}
    \emf = \int_C (\vec{E} + \vec{v} \times \vec{B}) \cdot \vec{dx}
\end{equation*}
This is again given by the line integral around $C$ of the Lorentz force on a particle of unit charge confined to $C$ (for which the perpendicular components of $\dvec{x}$ must agree with those of the curve velocity $\vec{v}$).
\begin{remark}
    Since we can choose any parameterisation $\lambda$, the curve velocity could have any number of tangential components (based on how the parameterisation moves around $C$), and so this is the strongest condition we could impose on $\dvec{x}$.
\end{remark}
\subsection{Lenz's Law}
Lenz's law states that the direction of the induced current is always such as to produce a magnetic field that opposes the change in flux that caused the emf.
\begin{example}
    Consider a circular wire in the $xy$-plane. If $B_z$ inside the loop increases in time then:
    \begin{equation*}
        \emf = -\frac{d\flux}{dt} < 0
    \end{equation*}
    which will induce a clockwise current around the loop ($I < 0$), which will generate a magnetic field $B_z < 0$ inside the loop. This avoid an unstable situation of positive feedback.
\end{example}
\section{Inductance and Magnetic Energy}
\subsection{Inductance}
\begin{definition}{Inductance}
    If a current $I$ around a circuit $C$ generates a magnetic field with flux $\flux$, then the \underline{inductance} of the circuit is defined by:
    \begin{equation*}
        L = \frac{\flux}{I}
    \end{equation*}
    and depends only on the geometry of the circuit (since $\flux$ is proportional to $I$).
\end{definition}
\begin{example}[Inductance of an Ideal Solenoid]
    Consider an ideal solenoid with cross-sectional area $A$ and $N$ turns per unit length. The magnetic field inside the solenoid is uniform:
    \begin{equation*}
        \vec{B} = \mu_0 N I \vec{e_z}.
    \end{equation*}
    Then the magnetic flux passing through this solenoid is $BA$ per turn, so the inductance per unit length is:
    \begin{equation*}
        l = \mu_0 N^2 A
    \end{equation*}
\end{example}
\begin{example}[Mutual inductance between two wires]
    Consider a thin wire $C_i$. Let the current around another thin wire $C_j$ be $I_j$. Then the flux through $C_i$ is:
    \begin{equation*}
        \flux_{ij} = L_{ij} I_j
    \end{equation*}
    where
    \begin{equation*}
        L_{ij} = \frac{\mu_0}{4\pi} \int_{C_i} \int_{C_j} \frac{\vec{dx_i} \cdot \vec{dx_j}}{|\vec{x_i} - \vec{x_j}|}
    \end{equation*}
    %TODO: Derive this with the magnetic vector potential.
\end{example}
\subsection{Changes in Current and Electromagnetic Energy}
When the current $I$ around a circuit $C$ is varied, an emf is induced:
\begin{equation*}
    \emf = -\frac{d\flux}{dt} = -L \frac{dI}{dt}
\end{equation*}
In a small time interval $\delta t$, a charge $\delta Q = I \delta t$ flows around the circuit and the work done on it by the Lorentz force is:
\begin{equation*}
    \delta W = \emf~\delta Q = -LI \frac{dI}{dt} \delta t
\end{equation*}
Then the rate at which work is done by the current on the electrogmagnetic field is:
\begin{equation*}
    -\frac{dW}{dt} = LI \frac{dI}{dt} = \frac{d}{dt} \left(\frac12 LI^2\right).
\end{equation*}
Consider reaching a magnetostatic state by building up the current from 0. Then the energy stored in this configuration is:
\begin{equation*}
    \int_0^t \frac{d}{dt} \left(\frac12 LI^2\right) dt = \frac12 LI^2
\end{equation*}
As with capacitance, we can also write:
\begin{align*}
    U &= \frac12 LI^2 \\
    &= \frac12 I \flux \\
    &= \frac12 I \int_C \vec{A} \cdot \vec{dx} \\
    &= \frac12 \int_{\R^3} \vec{J} \cdot \vec{A} dV
\end{align*}
which we can compare to the energy stored in an electric field:
\begin{equation*}
    U = \frac12 \int_{\R^3} \rho \Phi dV.
\end{equation*}
In electrostatics, we then converted this into an integral involving $|\vec{E}|^2$. In the case of magnetostatics, we apply \eqnref{eqnMTBC}:
\begin{equation*}
    U = \frac{1}{2\mu_0} \int_{\R^3} \left(\nabla \times \vec{B}\right)\cdot \vec{A} dV
\end{equation*}
then apply the identity $(\nabla \times \vec{B}) \cdot \vec{A} = \nabla \cdot (\vec{B} \times \vec{A}) + \vec{B} \cdot (\nabla \times \vec{A})$. Since we are taking the integral over all space, the first term is zero by the divergence theorem. This is because:
\begin{align*}
    |\vec{B}| &= O(|\vec{x}|^{-3}) \\
    |\vec{A}| &= O(|\vec{x}|^{-2})
\end{align*}
and so $\vec{B} \times \vec{A} = O(|\vec{x}|^{-5})$ which ``beats'' the increasing surface area which is $O(|\vec{x}|^2)$. Therefore the energy is:
\begin{equation*}
    U = \int_{\R^3} \frac{|\vec{B}|^2}{2\mu_0} dV
\end{equation*}
\section{Ohm's Law}
In a stationary conductor, Ohm's Law states that:
\begin{equation*}
    \vec{J} = \sigma \vec{E}
\end{equation*}
where $\sigma$ is the electrical conductivity. This is not a fundamental physical law, but is a constitutive relation (a macroscopic property of a material). We can invert this relation to get:
\begin{equation*}
    \vec{E} = \sigma^{-1} \vec{J}
\end{equation*}
where $\sigma^{-1}$ is the resistivity. It is often denoted $\rho$, but this is in conflict with the notation for electric charge density.
\begin{remark}
    In \underline{isotropic} materials, there is no preferred direction and $\sigma$ is a scalar. In \underline{anisotropic} materials, there is a preferred direction and instead $\sigma$ is a symmetric, second-order tensor. Therefore the resistivity is its inverse.
\end{remark}
A perfect conductor corresponds to the limit $\sigma \to \infty$ and so $\vec{E} = \vec0$. A perfect insulator corresponds to the limit $\sigma \to 0$, and so $\vec{J} = 0$.
\begin{example}[Straight wire]
    Consider a straight wire of length $l$, in the direction of the unit vector $\vec{n}$. Let the wire have uniform cross-sectional area $A$ and conductivity $\sigma$.

    Consider an electric field $\vec{E} = E\vec{n}$. Then Ohm's Law gives $\vec{J} = \sigma E\vec{n}$. The total current is the integral over a cross-sectional surface. Since this is uniform, $I = \sigma E A$. The potential difference along the wire (assuming finite length $L$) is the line integral of the electric field, $V = EL$. Therefore, the resistence $R$ is defined to be:
    \begin{equation*}
        R = \frac{L}{\sigma A}
    \end{equation*}
    and so $V = IR$, an alternate form for Ohm's law.
\end{example}
We can find the energy loss due to resistivity for a wire. We can consider \underline{Joule heating} (also known as Ohmic heating), which is the conversion of electromagnetic energy into heat at the rate $I^2 R$. We can see this as follows: suppose that the potential difference $V$ is maintained by an energy source. Then $VI = I^2 R$ is the rate at which the emf of the energy source does work to maintain the current $I$. By conservation of energy, this must be the energy lost to heating.
\section{Time-Dependent Electric Fields}
\subsection{Scalar and Vector Potentials}
In electrodynamics, we can no longer write the electric field as $\vec{E} = - \nabla \phi$. This is because $\nabla \times \vec{E} = \frac{d\vec{B}}{dt} \neq 0$. However, \eqnref{eqnMBD} still allows us to write $\vec{B} = \nabla \times \vec{A}$. Next, we rewrite \eqnref{eqnMEC} as:
\begin{equation*}
    \nabla \times \left(\vec{E} + \frac{\partial \vec{A}}{\partial t}\right) = 0
\end{equation*}
and so we can write this using a scalar potential $\phi$:
\begin{equation*}
    \vec{E} = -\nabla \Phi - \frac{\partial \vec{A}}{\partial t}.
\end{equation*}
We can still make a guage transformation:
\begin{align*}
    \tilde{\vec{A}} &= \vec{A} + \nabla \chi, & \tilde{\Phi} &= \Phi - \frac{\partial \chi}{\partial t}
\end{align*}
where $\chi(\vec{x}, t)$ is any scalar field. Then both $\vec{E}$ and $\vec{B}$ are unchanged.
\subsection{Displacement Current}
In magnetostatics, we used Ampere's Law:
\begin{equation*}
    \oint_C \vec{B} \cdot \vec{dx} = \mu_0 \int_S \vec{J} \cdot \vec{dS} = \mu_0 I
\end{equation*}
which was derived from \eqnref{eqnMTBC}: $\nabla \times \vec{B} = \mu_0 \vec{J}$. We now need to work with:
\begin{equation*}
    \nabla \times \vec{B} = \mu_0 \left(\vec{J} + \epsilon_0 \frac{\partial \vec{E}}{\partial t}\right)
    \tag{\ref{eqnMBC}}
\end{equation*}
and this extra term is called the \underline{displacement current}.

Why is it needed? Without this extra term, we would have $\nabla \cdot \vec{J} = 0$ which is our description of conservation of charge when $\rho$ is constant. Suppose now we place free particles of positive charge in some localised region. Repulive Coulomb forces cause the particles to separate, implying $\nabla \cdot \vec{J} > 0$. We have seen that charge conservation in the correct form,
\begin{equation*}
    \frac{\partial \rho}{\partial t} + \nabla \cdot \vec{J} = 0
\end{equation*}
follows from Maxwell's equations including the displacement current.
\section{Electromagnetic Waves}
\subsection{The Wave Equation}
Consider freely evolving electric and magnetic fields in a vacuum, in the absence of charges and currents.
\begin{align}
    \nabla \cdot \vec{E} &= 0 \label{eqnMWED} \\
    \nabla \cdot \vec{B} &= 0 \label{eqnMWBD} \\
    \nabla \times \vec{E} &= -\frac{\partial \vec{B}}{\partial t} \label{eqnMWEC} \\
    \nabla \times \vec{B} &= \mu_0 \epsilon_0 \frac{\partial \vec{E}}{\partial t} \label{eqnMWBC}
\end{align}
Eliminate $\vec{B}$ by taking the time derivative of \eqnref{eqnMWBC} and substituting \eqnref{eqnMWEC}:
\begin{align*}
    \mu_0 \epsilon_0 \frac{\partial^{2}\vec{E}}{\partial t^{2}} &= \frac{\partial}{\partial t} (\nabla \times \vec{B}) \\
    &= \nabla \times \frac{\partial \vec{B}}{\partial t} \\
    &= - \nabla \times (\nabla \times \vec{E}) \\
    &= \nabla^2 \vec{E} - \nabla (\nabla \cdot \vec{E}) \\
    &= \nabla^2 \vec{E} \text{ by \eqnref{eqnMWED}}.
\end{align*}
Alternatively, eliminating $\vec{E}$ by taking the time derivative of \eqnref{eqnMWEC} and substituting \eqnref{eqnMWBC}:
\begin{align*}
    \frac{\partial^{2}\vec{B}}{\partial t^{2}} &= - \frac{\partial}{\partial t} (\nabla \times \vec{E}) \\
    &= - \nabla \cdot \frac{\partial \vec{E}}{\partial t} \\
    &= - \frac{1}{\mu_0 \epsilon_0} \nabla \times (\nabla \times \vec{B}) \\
    &= \frac{1}{\mu_0 \epsilon_0} \nabla^2 \vec{B}
\end{align*}
Then we have:
\begin{align*}
    \frac{\partial^{2}\vec{E}}{\partial t^{2}} &= \frac{1}{\mu_0 \epsilon_0} \nabla^2 \vec{E} \\
    \frac{\partial^{2}\vec{B}}{\partial t^{2}} &= \frac{1}{\mu_0 \epsilon_0} \nabla^2 \vec{B}
\end{align*}
and so every Cartesian component of $\vec{E}$ and $\vec{B}$ satisfy:
\begin{equation*}
    \frac{\partial^{2}u}{\partial t^{2}} = c^2 \nabla^2 u
\end{equation*}
where $c$ is the speed of light, $c = \frac{1}{\sqrt{\mu_0 \epsilon_0}}$. We find that light is an electromagnetic wave involving oscillations of $\vec{E}$ and $\vec{B}$. Different values for the wavelength give different types of EM waves.
\subsection{Plane Electrogmagnetic Waves}
Consider a \underline{plane wave} in which $\vec{E}$ and $\vec{B}$ depend only on $x, t$ and not $y, z$ (that is, it depends only one one spatial direction and we choose it to be $x$). A simple example is:
\begin{equation*}
    \vec{E} = E(x, t) \vec{e_y}
\end{equation*}
where $E(x, t)$ satisfies the 1-dimensional wave equation:
\begin{equation*}
    \frac{\partial^{2}E}{\partial t^{2}} = c^2 \frac{\partial^{2}E}{\partial x^{2}}
\end{equation*}
which has general solution:
\begin{equation*}
    E(x, t) = f(x - ct) + g(x - ct)
\end{equation*}
which is a superposition of a wave travelling to the left and a wave travelling to the right. Now what is the corresponding magnetic field?
\begin{align*}
    \frac{\partial \vec{B}}{\partial t} &= - \nabla \times \vec{E} \\
    &= - \frac{\partial E}{\partial x} \vec{e_z} \\
    &= (-f'(x - ct) - g'(x - ct))
\end{align*}
then integrating with respect to $t$:
\begin{equation*}
    \vec{B}(x, t) = \frac1c\left(f(x - ct) - g(x - ct)\right) \vec{e_z}
\end{equation*}
which we could also have derived from \eqnref{eqnMWBC}. We also have constants here, but since they do not affect the wave we set them to 0.

Of particular importance is a monochromatic wave of a single angular frequency $\omega$. An example is:
\begin{align*}
    \vec{E} &= E_0 \cos(kx - \omega t) \vec{e_y} \\
    \vec{B} &= \frac{E_0}{c}\cos(kx - \omega t) \vec{e_z}
\end{align*}
where $E_0$ is the amplitude (constant), and $k$ is the wave number $\frac{\omega}{c}$.

\begin{remarks}
    \item The frequency is $\nu = \frac{\omega}{2\pi}$ and the period is $\frac1\nu = \frac{2\pi}{\omega}$.
    \item The angular frequency and wave number are related by the dispersion relation:
        \begin{equation*}
            \omega^2 = c^2 k^2 \iff \omega = \pm ck
        \end{equation*}
    \item The oscillations of $\vec{E}$ and $\vec{B}$ are in phase and orthogonal both to each other and the direction of motion. That is, the waves are transverse.
    \item Because Maxwell's Equations are linear, electromagnetic waves of different amplitudes, frequences and directions can be superposed.
\end{remarks}
\subsection{Polarisation}
A more general approach to plane EM waves is to look for solutions of Maxwell's Equations of the form:
\begin{align*}
    \vec{E} &= \Re(\vec{E_0} e^{i(\vec{k} \cdot \vec{x} - \omega t)}) \\
    \vec{B} &= \Re(\vec{B_0} e^{i(\vec{k} \cdot \vec{x} - \omega t)})
\end{align*}
where $\vec{E_0}$ and $\vec{B_0}$ are constant vectors in $\C^3$, the vector amplitudes, and $\vec{k} \in \R^3$ is the constant wavevector. $\omega \in \R$ is the angular frequency and the wave number is $|\vec{k}|$.

\begin{remark}
    If we replace the complex number in the brackets with its complex conjugate, the real part is unchanged. In this way, we can require $\omega > 0$ by correctly choosing the sign of $\vec{k}$.
\end{remark}

These field $\vec{E}$ and $\vec{B}$ must satisfy the wave equation. This reduces to the dispersion relation:
\begin{equation*}
    \omega^2 = c^2 |\vec{k}|^2
\end{equation*}
This individual Maxwell equations reduce to algebraic conditions:
\begin{align*}
    \nabla \cdot \vec{E} = 0 &\implies \vec{k} \cdot \vec{E_0} = 0 \\
    \nabla \cdot \vec{B} = 0 &\implies \vec{k} \cdot \vec{B_0} = 0 \\
    \nabla \times \vec{E} = -\frac{\partial \vec{B}}{\partial t} &\implies \vec{k} \times \vec{E_0} = \omega \vec{B_0} \\
    \nabla \times \vec{B} = \mu_0 \epsilon_0 \frac{\partial \vec{E}}{\partial t} &\implies \vec{k} \times \vec{B_0} = -\frac\omega{c^2} \vec{E_0}
\end{align*}
The fourth equation is redundant if we consider these equations alongside the dispersion relation. We combine the first and third equations with the dispersion relation to get:
\begin{align*}
    \vec{k} \times \vec{B_0} &= \frac{1}{\omega} \vec{k} \times (\vec{k} \times \vec{E_0}) \\
    &= -\frac{k^2}{\omega} \vec{E_0} \\
    &= - \frac{\omega^2}{c^2}\vec{E_0}
\end{align*}

Suppose $\vec{E_0}$ is real. Then $\vec{B_0}$ is also real and the vectors $\vec{k}, \vec{B_0}, \vec{E_0}$ are an orthogonal triad. Then $\vec{E}$ and $\vec{B}$ oscillate in fixed directions, which are perpendicular to each other and to the direction of propagation. This is similar to the one-dimensional wave considered previously and corresponds to a \underline{linearly polarised wave}. Polarisation describes the orientation of the electric field.

Now suppose that $\vec{E_0}$ is complex and of the form $\vec{a} - i \vec{b}$ where $\vec{a}$ and $\vec{b}$ are real. These must lie in the plane orthogonal to $\vec{k}$. We find:
\begin{equation*}
    \vec{E} = \vec{a} \cos(\vec{k} \cdot \vec{x} - \omega t) + \vec{b} \sin(\vec{k} \cdot \vec{x} - \omega t)
\end{equation*}
Assuming that $\vec{a}$ and $\vec{b}$ are linearly independent, this gives a direction for $\vec{E}$ that is not fixed, but rotates in the plane orthogonal to $\vec{k}$. In general it will trace out an ellipse specified by $\vec{a}$ and $\vec{b}$, and the electric field is said to be \underline{elliptically polarised}.

The special case of \underline{circular polarisation} occurs when $\vec{a}$ and $\vec{b}$ are orthogonal with the same magnitude. Then $|\vec{E}|$ traces out a circle. The polarisation is \underline{right-handed} if $\vec{b} = \uvec{k} \times \vec{a}$. Otherwise it is left-handed.
\subsection{Reflection from a Conductor}
Suppose a perfect conductor occupies the half-space $x > 0$. Let an incident wave be approaching the surface $x = 0$ normally with the form:
\begin{equation*}
    \vec{E_{inc}} = \Re\left(E_0 e^{i(kx - \omega t)}\right) \vec{e_y}
\end{equation*}
where we take $k > 0$, $\omega = ck > 0$ and $E_0 \in \R$. Because we are taking a perfect conductor, we have the boundary condition $\vec{E} = \vec0$ on $x = 0$. Since the tangential components of $\vec{E}$ are continuous at the interface, we require $E_y = 0$ at $x = 0$. The incident wave as seen above does not satisfy this boundary condition. In order to resolve this, we add the reflected wave and use the superposition principle:
\begin{equation*}
    \vec{E_{ref}} = \Re\left(-E_0 e^{i(-kx - \omega t)}\right) \vec{e_y}
\end{equation*}
which propagates in the $-x$ direction. Superposing the two waves:
\begin{align*}
    \vec{E} &= \vec{E_{inc}} + \vec{E_{ref}} \\
    &= \Re(2iE_0 \sin(kx) e^{-i\omega t}) \vec{e_y} \\
    &= 2E_0 \sin(kx) \sin(\omega t)\vec{e_y}
\end{align*}
which indeed solves the required boundary condition. We can interpret this as a standing wave with a node at $x = 0$.

The corresponding magnetic field is given by:
\begin{align*}
    \vec{B}&=\vec{B_{inc}} + \vec{B_{ref}} \\
    &= \Re\left[\frac{\vec{E_0} \times \vec{k}}{\omega}\left(e^{i(\vec{k} \cdot \vec{x} - \omega t)} + e^{i(-\vec{k} \cdot \vec{x} - \omega t)}\right)\right] \\
    &= \Re\left[\frac{E_0 k}{\omega} \vec{e_z}e^{-i\omega t}\left(e^{ikx} + e^{-ikx}\right)\right] \\
    &= \Re\left[\frac{E_0 k}{ck} \vec{e_z}e^{-i\omega t} \times 2\cos(kx)\right] \\
    &= \frac{2E_0}{c} \cos(\omega t) \cos(kx) \vec{e_z}
\end{align*}
which is a standing wave with an antipole at $x = 0$. Therefore, we have a surface discontinuity:
\begin{equation*}
    \left[\vec{B} \times \vec{n}\right] = \frac{2E_0}{c} \cos(\omega t)
\end{equation*}
and the surface charge is $\vec{K} = \frac{2E_0}{\mu_0 c} \cos(\omega t)$ on the conductor.
\section{Electromagnetic Energy}
\subsection{Equation of Energy Conservation}
In electrostatics, we found that the energy per unit volume in $\vec{E}$ was $\frac12 \epsilon_0 |\vec{E}|^2$. In magnetostatics, the energy per unit volume in $\vec{B}$ was $\frac12 \frac{1}{\mu_0} |\vec{B}|^2$.

We will now derive a general conservation law from Maxwell's Equations. Consider the rate of change of the electric and magnetic energies together, per unit volume:
\begin{align*}
    \frac{\partial}{\partial t}&\left(\frac{1}{2} \epsilon_0 |\vec{B}|^2 + \frac{1}{2} \frac{1}{\mu_0} |\vec{B}|^2\right) \\
    &= \vec{E} \cdot \left(\epsilon_0 \frac{\partial \vec{E}}{\partial t}\right) + \vec{B} \cdot \left(\frac{1}{\mu_0} \frac{\partial \vec{B}}{\partial t}\right) \\
    &= \vec{E} \cdot \left(\frac{1}{\mu_0} \nabla \times \vec{B} - \vec{J}\right) + \vec{B} \cdot \left(-\frac{1}{\mu_0} \nabla \times \vec{E}\right) \\
    &= -\frac{1}{\mu_0} \left(\vec{B} \cdot (\nabla \times \vec{E}) - \vec{E} \cdot (\nabla \times \vec{B})\right) - \vec{E} \cdot \vec{J} \\
    &= - \nabla \cdot \left(\frac{\vec{E} \times \vec{B}}{\mu_0}\right) - \vec{E} \cdot \vec{J}
\end{align*}
and this leads us to:
\begin{equation}
    \frac{\partial w}{\partial t} + \nabla \cdot \vec{S} = -\vec{E} \cdot \vec{J}
    \label{eqnConserveEnergy}
\end{equation}
where $w = \frac12 \epsilon_0 |\vec{E}|^2 + \frac12 \frac{1}{\mu_0} |\vec{B}|^2$ is the \underline{energy density} (energy per unit volume), $\vec{S} = \frac{\vec{E} \times \vec{B}}{\mu_0}$ is the \underline{energy flux density} (energy flux per unit area). The vector $\vec{S}$ is the \underline{Poynting vector}.

In the absence of currents, \eqnref{eqnConserveEnergy} becomes:
\begin{equation*}
    \frac{\partial w}{\partial t} + \nabla \cdot \vec{S} = 0
\end{equation*}
which looks like a more standard conservation law. We can integrate this over a fixed volume $V$ as we are used to (and use the divergence theorem):
\begin{equation*}
    \frac{\partial}{\partial t} \int_V w dV + \int_S \vec{S} \cdot \vec{dS} = 0
\end{equation*}
so the rate of change of energy in $V$ is equal to the flux out through $S$.

In the case of a nonzero RHS in \eqnref{eqnConserveEnergy}, $\vec{E} \cdot \vec{J}$ is the rate at which the electromagnetic field loses energy (per unit volume) via the Lorentz force.

Considering $N$ charged particles, with charges $q_i$ and positions $\vec{x_i}$,
\begin{align*}
    \rho &= \sum_{i=1}^{N} q_i \delta(\vec{x} - \vec{x_i}) \\
    \vec{J} &= \sum_{i=1}^{N} q_i \dvec{x_i} \delta(\vec{x} - \vec{x_i})
    \vec{F_i} &= q_i \left(\vec{E}(\vec{x_i}, t) + \dvec{x_i} \times \vec{B}(\vec{x_i}, t)\right)
\end{align*}
and so the rate of working is:
\begin{align*}
    \sum_{i=1}^{N} \vec{F_i} \cdot \dvec{x_i} &= \sum_{i=1}^{N} \vec{E}(\vec{x_i}, t) \cdot (q_i \vec{x_i}) \\
    &= \int_{\R^3} \vec{E} \cdot \vec{J}dV
\end{align*}
In a conductor, Ohm's Law is $\vec{J} = \sigma \vec{E}$. Therefore the rate of Joule heating is:
\begin{equation*}
    \int_{V} \vec{E} \cdot \vec{J} dV = \int_V \frac{|\vec{J}|^2}{\sigma} dV
\end{equation*}
which is positive definite and is the rate of Joule heating.
\subsection{Energy in Electromagnetic Waves}
Consider:
\begin{align*}
    \vec{E} &= \Re\left(\vec{E_0} e^{i(\vec{k} \cdot \vec{x} - \omega t)}\right) \\
    \vec{B} &= \Re\left(\vec{B_0} e^{i(\vec{k} \cdot \vec{x} - \omega t)}\right)
\end{align*}
and use the identity $\Re(A) \Re(B)  = \Re\left(\frac{AB + A\overline{B}}{2}\right)$:
\begin{align*}
    w &= \frac12 \epsilon_0 |\vec{E}|^2 + \frac{1}{2\mu_0} |\vec{B}|^2 \\
    &= \frac14 \epsilon_0 \Re\left((\vec{E_0} \cdot \vec{E_0}) e^{2 i (\vec{k} \cdot \vec{x} - \omega t)} + |\vec{E_0}|^2\right) \\
    &\quad+ \frac{1}{4\mu_0} \Re\left((\vec{B_0} \cdot \vec{B_0}) e^{2 i (\vec{k} \cdot \vec{x} - \omega t)} + |\vec{B_0}|^2\right)
\end{align*}
The oscillatory terms average to zero, leaving the average energy density:
\begin{equation*}
    \average{w} = \frac14 \epsilon_0 |\vec{E_0}|^2 + \frac1{4\mu_0} |\vec{B_0}|^2
\end{equation*}
Then since $\vec{B_0} = \frac1c \uvec{k} \times \vec{E_0}$ and $\uvec{k} \cdot \vec{E_0} = 0$, the two terms are equal and we can write:
\begin{equation*}
    \average{w} = \frac12 \epsilon_0 |\vec{E_0}|^2
\end{equation*}
Similarly, the Poynting vector is:
\begin{align*}
    \vec{S} &= \frac{\vec{E} \times \vec{B}}{\mu_0} \\
    &= \frac1{2\mu_0} \Re\left((\vec{E_0} \times \vec{B_0}) e^{2i (\vec{k} \cdot \vec{x} - \omega t)} + \vec{E_0} \times \overline{\vec{B_0}}\right)
\end{align*}
and we similarly take the average flux density:
\begin{align*}
    \average{\vec{S}} &= \frac{|\vec{E_0}|^2}{2\mu_0 c} \uvec{k} \\
    &= \frac12 \epsilon_0 |\vec{E_0}|^2 c \uvec{k} \\
    &= \average{w} c \uvec{k}
\end{align*}
Which implies that the electromagnetic wave transports energy at speed $c$ in the direction $\uvec{k}$.
\end{document}