\documentclass[../Main.tex]{subfiles}

\begin{document}
\section{Review of Relativity}
\subsection{Lorentz Transformation}
In the special theory of relativity we use spacetime coordinates:
\begin{equation*}
    X = \begin{pmatrix}ct \\ x \\ y \\ z\end{pmatrix} = \begin{pmatrix}ct \\ \vec{x}\end{pmatrix}
\end{equation*}
and reference the coordinates of $X$ as $X^\mu$. We use Greek letters for the components of 4-vectors. These run from $0$ to $3$. Roman indices will run from $1$ to $3$ to only reference spatial dimensions.

$X$ is the position 4-vector. Under a Lorentz transformation (LT) from frame $S$ to $S'$, it transforms according to
\begin{equation}
    X' = \Lambda X
    \label{eqnLorentzTransform}
\end{equation}
In index notation, we would find $X^{\prime \mu} = \indx{\Lambda}{\mu}{\nu} X^\nu$.

$\Lambda$ is any matrix in the Lorentz group. We are particularly interested in the Lorentz boost, where $S'$ moves relative to $S$ with speed $v < c$ in the $x$ direction.
\begin{equation*}
    \Lambda =
    \begin{pmatrix}
        \gamma & -\gamma \beta & 0 & 0 \\
        -\gamma\beta & \gamma & 0 & 0 \\
        0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 1
    \end{pmatrix}
\end{equation*}
where $\beta = \frac{v}{c} \in (-1, 1)$, and:
\begin{equation*}
    \gamma = \frac{1}{\sqrt{1 - \beta^2}}
\end{equation*}
is the \underline{Lorentz factor}.
\subsection{Four-Vectors}
Four-vectors will be denoted with captial Roman letters. Any 4-vector $V$ transforms according to \eqnref{eqnLorentzTransform}, $V' = \Lambda V$. We use the convention that any such vector can be written as:
\begin{equation*}
    V = \begin{pmatrix}V^0 \\ \vec{v}\end{pmatrix}
\end{equation*}
The pseudo-inner product of 4-vectors is:
\begin{equation*}
    V \cdot W = \vec{v} \cdot \vec{w} - V^0 W^0 = \eta_{\mu \nu} V^\mu W^\nu
\end{equation*}
where $\eta$ is the \underline{Minkowski metric}, and it is an isotropic tensor. It has components:
\begin{equation*}
    \eta_{\mu\nu} = \begin{pmatrix}
        -1 & 0 & 0 & 0 \\
        0 & 1 & 0 & 0 \\
        0 & 0 & 1 & 0 \\
        0 & 0 & 0 & 1
    \end{pmatrix}
    = \operatorname{diag}(-1, 1, 1, 1)
\end{equation*}
in any intertial frame.

Then the inner product is invariant under a Lorentz transformation. We see this:
\begin{align*}
    V' \cdot W' &= \eta_{\mu\nu}' V^{\prime \mu} W^{\prime \nu} \\
    &= \eta_{\mu \nu} \indx{\Lambda}{\mu}{\rho} V^\rho = \indx{\Lambda}{\nu}{\rho} W^\sigma \\
    &= \eta_{\rho \sigma} V^\rho W^\sigma \\
    &= V \cdot W
\end{align*}
because the LT matrix satisfies:
\begin{equation*}
    \eta_{\mu \nu} \indx{\Lambda}{\mu}{\rho} \indx{\Lambda}{\nu}{\sigma} = \eta_{\rho \sigma}
\end{equation*}
or, in matrix notation,
\begin{equation*}
    \Lambda^T \eta \Lambda = \eta
\end{equation*}
which implies $\det(\Lambda) = \pm 1$.

We can define the square of a $4$-vector to be $V \cdot V$:
\begin{equation*}
    V \cdot V = -V^0 V^0 + |\vec{v}|^2
\end{equation*}
and we say:
\begin{itemize}
    \item $V$ is \underline{timelike} if $V \cdot V < 0$,
    \item $V$ is \underline{spacelike} if $V \cdot V > 0$,
    \item $V$ is \underline{null} or \underline{lightlike} if $V \cdot V = 0$.
\end{itemize}
The invariant spacetime interval between two events (an event is a point and a time):
\begin{equation*}
    ds^2 = dX \cdot dX = -c^2 dt^2 + dx^2 + dy^2 + dz^2
\end{equation*}

\begin{definition}{4-velocity}
    The \underline{4-velocity} of a massive particle is:
    \begin{equation*}
        U = \frac{dX}{d\tau}
    \end{equation*}
\end{definition}
where $\tau$ is the proper time (time experienced by the particle).

Since $X = (ct, \vec{x})$,
\begin{align*}
    U &= \frac{dt}{d\tau} \begin{pmatrix}c \\ \vec{u}\end{pmatrix} \\
    &= \gamma_{\vec{u}} \begin{pmatrix}c \\ \vec{u}\end{pmatrix}    
\end{align*}
where $\vec{u}$ is the normal 3-velocity and $\gamma_{\vec{u}}$ is the Lorentz factor for the motion:
\begin{equation*}
    \gamma_{\vec{u}} = \frac{1}{1 - \frac{|\vec{x}|^2}{c^2}}
\end{equation*}
As for all 4-vectors, the dot product is invariant:
\begin{align*}
    U \cdot U &= \left(\frac{dt}{d\tau}\right)^2 (-c^2 + |\vec{u}|^2) \\
    &= -c^2
\end{align*}
Therefore this is invariant and timelike.

\begin{definition}{4-momentum}
    The \underline{4-momentum} of a massive particle with rest mass $m$ is:
    \begin{equation*}
        P = mU = m\gamma_{\vec{u}} \begin{pmatrix}c \\ \vec{u}\end{pmatrix} = \begin{pmatrix}\frac{E}{c} \\ \vec{p}\end{pmatrix}
    \end{equation*}
    where $\vec{p}$ is the 3-momentum.
\end{definition}
We know that the inner product is invariant:
\begin{align*}
    P \cdot P &= -\left(\frac{E}{c}\right)^2 + |\vec{p}|^2 \\
    &= -m^2 c^2
\end{align*}
which is consistent with $E^2 = m^2 c^4 + |\vec{p}|^2 c^2$.

In the Newtonian limit $v \ll c$,
\begin{align*}
    E &= m \gamma_{\vec{u}} c^2 \\
    &\approx m\left(1 + \frac12 \frac{|\vec{u}|^2}{c^2}\right) c^2 \\
    &= mc^2 + \frac12 m |\vec{u}|^2
\end{align*}
and in the rest frame where $\vec{u} = \vec0$, we simply have the rest energy $E = mc^2$.

Massless particles such as photons travel at the speed of light (in any inertial frame). They have $E = |\vec{p}|c$ and their 4-velocity is null.
\section{Covectors and Tensors}
We understand a 4-vector to be some $X$ in the relevant vector space, with components $X^\mu$ (superscript indices).
\begin{definition}{Covector}
    A \underline{covector} is an element $Y$ of the dual space, with components $Y_\mu$ (subscript indices).
\end{definition}
Define:
\begin{equation*}
    Y_\mu = \frac{\partial f}{\partial X^\mu}
\end{equation*}
and this is the derivative of a scalar field $f(X)$ in spacetime. This is a covector, and so we want to know how this transforms under a Lorentz transformation. By the chain rule,
\begin{align}
    \frac{\partial f}{\partial X^{\prime \mu}} &= \frac{\partial X^\nu}{\partial X^{\prime \mu}} \frac{\partial f}{\partial X^\nu} \nonumber \\
    Y_\mu' &= \indx{(\Lambda^{-1})}{\nu}{\mu} Y_\nu \label{eqnLorentzTransformDerivative}
\end{align}
where $(\Lambda^{-1})$ is the inverse Lorentz transformation matrix. \Eqnref{eqnLorentzTransformDerivative} is the general transformation law for covectors. Written without the indices,
\begin{equation*}
    Y' = \Lambda^{-1} Y
\end{equation*}
\begin{definition}{Tensor}
    A \underline{(4-)tensor} $T$ of type $(r, s)$ is an object whose $4^{r+s}$ elements transform as:
    \begin{equation}
        \indx{T}{\prime \mu_1 \cdots \mu_r}{\nu_1 \cdots \nu_s} = \indx{\Lambda}{\mu_1}{\rho_1} \cdots \indx{\Lambda}{\mu_r}{\rho_r} \indx{(\Lambda^{-1})}{\sigma_1}{\nu_1} \cdots \indx{(\Lambda^{-1})}{\sigma_s}{\nu_s} \indx{T}{\rho_1 \cdots \rho_r}{\sigma_1\cdots\sigma_s}
        \label{eqnLorentzTransformTensor}
    \end{equation}
    That is, a LT matrix for each superscript index and an inverse LT matrix for each subscript index.
\end{definition}
\begin{remarks}
    \item Any ordering of the up and down indices is possible. For example, a tensor of type $(1, 1)$, $A_\mu\hphantom{}^\nu$
    \item A $(1, 0)$ tensor is a 4-vector.
    \item A $(0, 1)$ tensor is a 4-covector.
    \item A $(0, 0)$ tensor is a scalar (Lorentz invariant).
\end{remarks}
\begin{propositions}{
        
        \label{propsTensorProps}
    }
    \item A linear combination of $(r, s)$ tensors is also an $(r, s)$ tensor. \label{propTensorLinearCombo}
    \item The \underline{outer product} of tensors of type $(r_1, s_1)$ and $(r_1, s_2)$ gives a tensor of type $(r_1 + r_2, s_1 + s_2)$. \label{propTensorOuterProd}
    \item The \underline{derivative} of a tensor field of type $(r, s)$ is another tensor field of type $(r, s+1)$.
        \begin{equation*}
            A_\mu\indx{}{\nu \rho}{\sigma} = \partial_\mu \indx{B}{\nu \rho}{\sigma}
        \end{equation*}
        where $\partial_\mu = \frac{\partial}{\partial X^\mu}$. \label{propTensorDerivative}
    \item The \underline{contraction} of an $(r, s)$ tensor on a pair of its indices produces an $(r-1, s-1)$ tensor. This only works when an up index is contracted with a down index. \label{propTensorContract}
\end{propositions}
The summation convention on spacetime requires any repeated index to appear once up and once down. This is because a contraction is an application of an element of the dual space to an element of the vector space. For example, the invalid expression $X^\mu X^\mu$ is not Lorentz invariant and so is not valid as a scalar.

\begin{definition}{Isotropic}
    A tensor is \underline{isotropic} if its components are the same in every inertial frame.
\end{definition}
\begin{remark}
    An example of this is the Minkowski metric $\eta_{\mu \nu}$. This is an isotropic tensor of type $(0, 2)$. The inverse of the Minkowski metric is $\eta^{\mu\nu}$ which is an isotropic tensor of type $(2, 0)$. Both have the same components, but they have different roles.
\end{remark}
The Kronecker Delta is an isotropic tensor of type $(1, 1)$. $\delta = diag(1,1, 1, 1)$.

The Levi-Civita epsilon symbol is defined by:
\begin{equation*}
    \epsilon_{\mu\nu\rho\sigma} =
    \begin{cases}
        1 & \text{if $(\mu, \nu, \rho, \sigma)$ is an even permutation of $(0,1,2,3)$} \\
        -1 & \text{if $(\mu, \nu, \rho, \sigma)$ is an odd permutation of $(0,1,2,3)$} \\
        0 & \text{otherwise}
    \end{cases}
\end{equation*}
and it is an isotropic pseudo-tensor of type $(0, 4)$. A pseudo-tensor satisfies the same transformation law as a tensor but with an extra factor of $\det(\Lambda)$.
\begin{definition}{Tensor equation}
    A \underline{tensor equation} is an equation in which every term is a tensor of the same type (transforms identically under a LT). Such an equation is said to be \underline{covariant} and is valid in any inertial frame.
\end{definition}
\section{Vector Calculus in Spacetime}
\subsection{Differential Operators}
The spacetime derivative of a scalar field $f(X)$ is a covector field:
\begin{equation*}
    \partial_\mu f = \begin{pmatrix}\frac1c \frac{\partial f}{\partial t} \\ \nabla f\end{pmatrix}
\end{equation*}
The spacetime divergence of a 4-vector field $V(X) = (V^0, \vec{v})^T$ is a scalar field:
\begin{equation*}
    \partial_\mu V^\mu(X) = \frac1c \frac{\partial V^0(X)}{\partial t} + \nabla \cdot \vec{v}
\end{equation*}
Curl is slightly more involved. Indeed, curl is only defined in $3$ dimensions. It involves the antisymmetric part of the derivatives of a vector field. In spacetime, we consider the antisymmetric derivative of a covector field $Y(X)$, which produces a tensor field of type $(0, 2)$.
\begin{equation*}
    \partial_\mu Y_\nu - \partial_\nu Y_\mu
\end{equation*}

The generalisation of the Laplacian operator $\nabla^2$ is the \underline{d'Alambertian} operator:
\begin{equation*}
    \Box = \eta^{\mu \nu} \partial_\mu \partial_\nu = -\frac{1}{c^2} \frac{\partial^{2}}{\partial t^{2}} + \nabla^2
\end{equation*}

The wave equation for a scalar field $f(X)$ is:
\begin{equation*}
    \Box f = 0
\end{equation*}
but this is a scalar equation only for waves that travel at the speed of light.
\subsection{Raising and Lowering}
The metric tensor provides an isomorphism between 4-vectors (with up indices) and covectors (with down indices). A vector $V$ can be related to a covector $W$ as:
\begin{equation*}
    V^\mu = \eta^{\mu \nu} W_\nu,\qquad W_\mu = \eta_{\mu \nu} V^\nu
\end{equation*}
these operations are known as \underline{raising and lowering indices}. Considering components, these raising and lowering operations will change the sign of the $0$th (time) component, while leaving the spatial components unchanged.

For example, the 4-velocity is lowered as: $U_\mu = \gamma_{\vec{u}} (-c, \vec{u})$.

A derivative is raised as: $\partial^\mu f = \left(-\frac{1}{c} \frac{\partial f}{\partial t}, \nabla f\right)$.
The inner product of 4-vectors can now be written as:
\begin{align*}
    V \cdot W &= \eta_{\mu\nu} = V^\mu W^\nu \\
    &= V_\nu W^\nu \\
    &= V^\mu W_\mu \\
    &= \eta^{\mu \nu} V_\nu W_\mu
\end{align*}
Any tensor index can be raised or lowered in the same way. For example, a tensor $T^{\mu \nu}$ can have its indices lowered to produce:
\begin{align*}
    \indxd{T}{\mu}{\nu} &= \eta_{\mu \rho} T^{\rho \nu} \\
    \indx{T}{\mu}{\nu} &= \eta_{\nu \rho}T^{\mu \rho}
\end{align*}
%TODO EXERCISE: show that raising an index of $\eta$ gives the Kronecker Delta: $\eta^\mu_\nu = \delta^\mu_\nu$.
A tensor may be symmetric or antisymmetric on a pair of indices if they are both down or both up.
%TODO: Show that this is preserved if both indices are raised/lowered.
\section{Charge Conservation and Motivating Examples}
Recall the equation for charge conservation in the non-relativistic setting:
\begin{equation*}
    \frac{\partial \rho}{\partial t} + \nabla \cdot \vec{J} = 0 \tag{\ref{eqnChargeConserve}}
\end{equation*}
We see that this is naturally written in the form of a spacetime divergence:
\begin{equation}
    \partial_\mu J^\mu = 0
    \label{eqnChargeConserveRelativ}
\end{equation}
where here $J$ is a new 4-vector called the \underline{charge-current} density or the 4-current density:
\begin{equation*}
    J = \begin{pmatrix}c\rho \\ \vec{J}\end{pmatrix}
\end{equation*}
\begin{example}
    Consider a Lorentz boost of speed $v$ in the $x$ direction.

    In $S$, let there be a uniform distribution of charged particles at rest, each with charge $q$ and with $n$ particles per unit volume. Then $\rho = nq$ and:
    \begin{equation*}
        J = \begin{pmatrix}nqc \\ \vec{0}\end{pmatrix}
    \end{equation*}
    In $S'$, we would expect to see particles moving with $3$-velocity $-v\vec{e_x}$. Length contraction means that the density of particles increases to $\gamma n$. We would therefore expect:
    \begin{equation*}
        J' = \begin{pmatrix}\gamma n q c \\ -\gamma n q v\vec{e_x}\end{pmatrix}
    \end{equation*}
    and by performing the Lorentz transformation we see that this is correct:
    \begin{equation*}
        \begin{pmatrix}
            \gamma & -\gamma v / c & 0 & 0 \\
            -\gamma v / c & \gamma & 0 & 0 \\
            0 & 0 & 1 & 0 \\
            0 & 0 & 0 & 1
        \end{pmatrix}
        \begin{pmatrix}nqc \\ \vec{0}\end{pmatrix}
        = \begin{pmatrix}\gamma n q c \\ -\gamma n q v \\ 0 \\ 0\end{pmatrix}
    \end{equation*}
\end{example}
\begin{example}
    Consider, in a frame $S$, a wire carrying particles of charge $q$ (with density $n+$) moving along the wire with 3-velocity $u\vec{e_x}$. Let there also be particles of charge $-q$, with density $n_- = n_+$ moving with 3-velocity $-u\vec{e_x}$. Therefore, the overall charge density is $\rho = 0$. The net current density is:
    \begin{equation*}
        \vec{J} = n_+ qu\vec{e_x} + n_0 (-q) (-u) \vec{e_x} = 2n_+ q u \vec{e_x}
    \end{equation*}
    Then the wire is uncharged but carries a current. This will have no electric field, and a magnetic field with field lines travelling in circles around the line of axis of the wire.

    In $S'$, which moves at speed $v$ in the $x$ direction relative to $S$, the positive and negative charged particles will have different velocities in the $x$ direction:
    \begin{equation*}
        u_{\pm}' = \frac{\pm u - v}{1 \mp \frac{uv}{c^2}}
    \end{equation*}
    and therefore we will have different charge densities for the positive and negative charges. This means we will have $\rho' \neq 0$, leading to a non-zero electric field in frame $S'$. We can get the complementary result ($\vec{B} = 0$ in $S$ but $\vec{B} \neq 0$ in $S'$) by considering the previous example.
    
    These examples show that $\vec{E}$ and $\vec{B}$ must be linked in relativistic electromagnetism. We will see in the next section that indeed they form the electromagnetic tensor.
\end{example}
\section{Electromagnetic Tensor}
The relativistic equation of motion for a massive particle is:
\begin{equation}
    \frac{\partial P^\mu}{\partial \tau} = f^\mu
    \label{eqnMotionRelativ}
\end{equation}
where $f^\mu$ is the 4-force.

Since $P \cdot P = -m^2 c^2$ is constant, $f^\mu$ must satisfy $f \cdot P = 0$. This is because,
\begin{align*}
    0 &= \frac{\partial}{\partial \tau}(P \cdot P) \\
    &= \frac{\partial }{\partial \tau} \left(\eta_{\mu \nu} P^\mu P^\nu\right) \\
    &= \eta_{\mu \nu} \left(\frac{\partial P^\mu}{\partial \tau} P^\nu + P^\mu \frac{\partial P^\nu}{\partial \tau}\right) \\
    &= \eta_{\mu \nu} \left(f^\mu p^\nu + p^\mu f^\nu\right)\\
    &= 2\eta_{\mu \nu} f^\mu P^\nu = 2f \cdot P
\end{align*}
So now, how can we express the Lorentz force as a 4-vector? It should be proportional to $q$ and linear in $U^\mu$. We assume:
\begin{equation}
    f_\mu = qF_{\mu \nu} U^\nu
    \label{eqnForceGuess}
\end{equation}
for some $(0, 2)$ tensor $F_{\mu\nu}$. Our condition of $f \cdot P = 0$ gives:
\begin{equation*}
    qm F_{\mu \nu} U^\mu U^\nu = 0
\end{equation*}
which is satisfied for all $U$ if $f$ is antisymmetric.

The spatial components of \eqnref{eqnForceGuess} are:
\begin{align*}
    f_i &= q\left(f_{i0} U^0 + F_{ij} U^j\right) \\
    &= \gamma_{\vec{u}} q \left(F_{i0} c + F_{ij} u_j\right)
\end{align*}
and we would like this to be equal to our existing expression for the force:
\begin{equation*}
    \gamma_{\vec{u}} q(E_i + \epsilon_{ijk} u_j B_k)
\end{equation*}
where the Lorentz factor appears because we are differentiating with respect to proper time.

We therefore require:
\begin{equation*}
    F_{i0} = \frac{E_i}{c},\quad F_{ij} = \epsilon_{ijk} B_k
\end{equation*}
and by antisymmetry, $F_{00} = 0, F_{0i} = -\frac{E_i}{c}$. Writing this as a matrix:
\begin{equation}
    F_{\mu\nu} =
    \begin{pmatrix}
        0 & -E_x / c & -E_y / c & -E_z / c \\
        E_x / c & 0 & B_z & -B_y \\
        E_y / c & -B_z & 0 & B_x \\
        E_z / c & B_y & -B_x & 0
    \end{pmatrix}
    \label{eqnEMTensor}
\end{equation}
and this is our expression for the \underline{electromagnetic field tensor}. Note that neither $\vec{B}$ nor $\vec{E}$ is the spatial part of a 4-vector. However, for a particle with $4$-velocity $U$, we can calculate the covector:
\begin{equation*}
    F_{\mu \nu} U^\nu
\end{equation*}
and, in the rest frame of the particle, we find that it has components:
\begin{equation*}
    \begin{pmatrix}0 \\ \vec{E}\end{pmatrix}
\end{equation*}
By construction, the spatial components of \eqnref{eqnForceGuess} give the relativistic equation of motion:
\begin{align*}
    \frac{\partial \vec{p}}{\partial \tau} &= \frac{\partial (m\gamma_{\vec{u}} \vec{u})}{\partial \tau} \gamma_{\vec{u}} q(\vec{E} + \vec{u} \times \vec{B}) \\
    \frac{\partial \vec{p}}{\partial t} &= q(\vec{E} + \vec{u} \times \vec{B}) = \vec{F}
\end{align*}
where $\vec{F}$ is the normal (3-)Lorentz force.

The time component of \eqnref{eqnForceGuess} is now:
\begin{align*}
    \frac{\partial P^0}{\partial \tau} &= f^0 = -f_0 \\
    &= -q F_{0i} U^i \\
    &= q \frac{E_i}{c} \gamma_{\vec{u}} u_i
\end{align*}
which is equivalent to the energy equation (recall that the $0$th component of $P$ is the energy over $c$):
\begin{equation*}
    \frac{\partial E}{\partial \tau} = q \gamma_{\vec{u}} \vec{E} \cdot \vec{u}
\end{equation*}
Here scalar $E$ is the energy, and vector $\vec{E}$ is the electric field. To avoid confusion, sometimes we write $\mathcal{E}$ for the energy. (But this conflicts with our notation for EMF!)

We can also write:
\begin{equation*}
    \frac{\partial E}{\partial t} = q \vec{E} \cdot \vec{u} = \vec{F} \cdot \vec{u}
\end{equation*}
\section{The 4-Potential}
We previously wrote:
\begin{equation*}
    \vec{E} = -\nabla \Phi - \frac{\partial \vec{A}}{\partial t},\quad \vec{B} = \nabla \times \vec{A}
\end{equation*}
in terms of scalar and vector potentials. In relativity, we see that $\vec{E}$ and $\vec{B}$ are combined into a tensor, so we want to relate $F_{\mu \nu}$ to potentials like this.
\begin{align*}
    F_{i0} &= \frac{E_i}{c} = -\frac1c \frac{\partial \Phi}{\partial x_i} - \frac1c \frac{\partial A_i}{\partial t} \\
    &= \partial_i \left(-\frac{\Phi}{c}\right) - \partial_0 A_i \\
    F_{ij} &= \epsilon_{ijk} B_k = \epsilon_{ijk} \epsilon_{klm} \frac{\partial A_m}{\partial x_l} \\
    &= \partial_i A_j - \partial_j A_i
\end{align*}
and so we interpret $F_{\mu \nu}$ as the antisymmetric derivative:
\begin{equation}
    F_{\mu \nu} = \frac{\partial A_\nu}{\partial X_\mu} - \frac{\partial A_\mu}{\partial X_\nu}
    \label{eqnForcePotential}
\end{equation}
of the covector:
\begin{equation*}
    A_\mu = \left(-\frac{\Phi}{c}, \vec{A}\right).
\end{equation*}
Raising the index gives the 4-vector potential:
\begin{equation}
    A^\mu =
    \begin{pmatrix}\frac{\Phi}{c} \\ \vec{A}\end{pmatrix}
    \label{eqnFourPotential}
\end{equation}
A gauge transformation:
\begin{equation*}
    \tilde{\vec{A}} = \vec{A} + \nabla \chi,\quad \tilde{\Phi} = \Phi - \frac{\partial \chi}{\partial t}
\end{equation*}
corresponds to replacing $A_\mu$ with:
\begin{equation*}
    \tilde{A}_{\mu} = A_\mu + \partial_\mu \chi
\end{equation*}
that is, adding the covector
\begin{equation*}
    \left(\frac1c \frac{\partial \chi}{\partial t}, \nabla \chi\right)
\end{equation*}
to $A_\mu$. This leaves $F_{\mu \nu}$ unaffected.

The spacetime equivalent of the Coulomb gauge is the Lorentz gauge, in which:
\begin{equation*}
    \partial_\mu A^\mu = 0
\end{equation*}
We can make a gauge transformation to get to Lorentz gauge by choosing $\chi$ to be the solution of the forced wave equation:
\begin{equation*}
    -\Box \chi = \partial_\mu A^\mu
\end{equation*}
\section{Lorentz Transformation of the Electromagnetic Tensor}
Since $F_{\mu \nu}$ is a $(0, 2)$ tensor, it transforms as:
\begin{equation*}
    F_{\mu \nu}' = \indx{(\Lambda^{-1})}{\rho}{\mu} \indx{(\Lambda^{-1})}{\sigma}{\nu} F_{\rho \sigma}
\end{equation*}
In matrix notation, $F' = (\Lambda^{-1})^T F \Lambda^{-1}$.

For a Lorentz boost of $v$ along $x$,
\begin{align*}
    F_{01}' &= \gamma^2 (F_{01} + \beta^2 F_{10} + \beta(F_{00} + F_{11})) \\
    &= \gamma^2 (1 - \beta^2) F_{01} \\
    &= F_{01}
\end{align*}
so the electric field component along the direction of the boost is unchanged. We find:
\begin{align*}
    F_{02}' &= \gamma (F_{02} + \beta F_{12}) & F_{03}' &= \gamma(F_{03} + \beta F_{13}) \\
    F_{12}' &= \gamma (F_{12} + \beta F_{02}) & F_{13}' &= \gamma(F_{13} + \beta F_{03}) \\
    F_{23}' &= F_{23}
\end{align*}
from which we deduce:
\begin{align*}
    E_x' &= E_x & B_x' &= B_x \\
    E_y' &= \gamma (E_y - v B_z) & B_y' &= \gamma \left(B_y + \frac{v}{c^2} E_z\right) \\
    E_z' &= \gamma (E_z + v B_y) & B_z' &= \gamma \left(B_z - \frac{v}{c^2} E_y\right)
\end{align*}
More generally, if $\vec{v}$ is the velocity of $S'$ with respect to $S$,
\begin{align*}
    \vec{E_\parallel}' &= \vec{E_\parallel} & \vec{B_\parallel}' &= \vec{B_\parallel} \\
    \vec{E_\perp}' &= \gamma (\vec{E_\perp} + \vec{v} \times \vec{B_\perp}) & \vec{B_\perp}' &= \gamma \left(\vec{B_\perp} - \frac{1}{c^2} \vec{v} \times \vec{E_\perp}\right)
\end{align*}
by considering the components perpendicular and parallel to $\vec{v}$.

Consider the tensor formed by raising both indices of the electromagnetic tensor:
\begin{equation*}
    F^{\mu \nu} =
    \begin{pmatrix}
        0 & E_x / c & E_y / c & E_z / c \\
        -E_x / c & 0 & B_z & -B_y \\
        -E_y / c & -B_z & 0 & B_x \\
        -E_z / c & B_y & -B_x & 0
    \end{pmatrix}
\end{equation*}
We can use this to find two Lorentz-invariant combinations of $\vec{E}$ and $\vec{B}$:
\begin{align*}
    F_{\mu \nu} F^{\mu \nu} &= -\frac2{c^2} \left(|\vec{E}|^2 - c^2 |\vec{B}|^2\right) \\
    \epsilon_{\mu \nu \rho \sigma} F^{\mu \nu} F^{\rho \sigma} = \frac8c \vec{E} \cdot \vec{B}
\end{align*}
\begin{remark}
    The second combination is using the dual of the electromagnetic tensor
\end{remark}
So we conclude that $|\vec{E}|^2 - c^2 |\vec{B}|^2$ is a scalar, and $\vec{E} \cdot \vec{B}$ is a pseudoscalar (which changes sign under time reversal and spatial reflections). If $\vec{E}$ and $\vec{B}$ are perpendicular in one frame, they are perpendicular in all frames.
\section{Maxwell's Equations}
Recall Maxwell's Equations:
\begin{align*}
    \nabla \cdot \vec{E} &= \frac{\rho}{\epsilon_0} \tag{\ref{eqnMED}}\\
    \nabla \cdot \vec{B} &= 0 \tag{\ref{eqnMBD}} \\
    \nabla \times \vec{E} &= -\frac{\partial \vec{B}}{\partial t} \tag{\ref{eqnMEC}} \\
    \nabla \times \vec{B} &= \mu_0 \left(\vec{J} + \epsilon_0 \frac{\partial \vec{E}}{\partial t}\right) \tag{\ref{eqnMBC}}
\end{align*}
We must use the relations:
\begin{align*}
    F^{0i} &= -F^{i0} = \frac{E_i}{c},\quad F^{ij} = \epsilon_{ijk} B_k \\
    J^0 &= \rho c,\quad J^{i} = J_i,\quad c^2 = \frac1{\mu_0 \epsilon_0}
\end{align*}
Then \eqnref{eqnMED} becomes:
\begin{equation*}
    c \partial_i F^{0i} = \frac{J_0}{\epsilon_0 c} \implies \partial_i F^{0i} = \mu_0 J^0
\end{equation*}
and \eqnref{eqnMBC} becomes:
\begin{align*}
    \epsilon_{ijk} \partial_j B_k &= \mu_o J_i + \frac1c \partial_0 E_i \\
    \implies \partial_0 F^{i0} &+ \partial_j F^{ij} = \mu_o J^i
\end{align*}
and these are the components of the 4-vector equation:
\begin{equation}
    \partial_\nu F^{\mu \nu} = \mu_0 J^\mu
    \label{eqnMR1}
\end{equation}
Next, considering $B_i = \frac12 \epsilon_{ijk} F_{jk}$, \eqnref{eqnMBD} becomes:
\begin{align*}
    0 &= \partial_i B_i \\
    &= \frac12 \epsilon_{ijk} \partial_i F_{jk} \\
    &= \partial_1 F_{23} + \partial_2 F_{31} + \partial_3 F_{12}
\end{align*}
which is equivalent to:
\begin{equation}
    \partial_i F_{jk} + \partial_j F_{ji} + \partial_k F_{ij} = 0
    \label{eqnMaxwellIntermediate}
\end{equation}
because $F_{ij}$ is antisymmetric. \Eqnref{eqnMEC} becomes:
\begin{align*}
    \epsilon_{ijk} \partial_j E_k &= -c \partial_0 B_i \\
    \implies c \epsilon_{ijk} \partial_j F_{k0} &+ \frac{c}{2} \epsilon_{ijk} \partial_0 F_{jk} = 0
\end{align*}
then if we multiply by $\epsilon_{ilm}$ and divide by $c$:
\begin{equation*}
    \partial_l F_{m0} - \partial_m F_{l0} + \partial_0 F_{lm} = 0
\end{equation*}
which, by using antisymmetry and re-labelling, gives:
\begin{equation*}
    \partial_i F_{j0} + \partial_j F_{0i} + \partial_0 F_{ij} = 0
\end{equation*}
which, when combined with our earlier \eqnref{eqnMaxwellIntermediate} to:
\begin{equation}
    \partial_\mu F_{\nu \rho} + \partial_\nu F_{\rho \mu} + \partial_\rho F_{\mu \nu}
    \label{eqnMR2Expanded}
\end{equation}
and can be alternatively written as:
\begin{equation}
    \epsilon^{\mu \nu \rho \sigma} \partial_\nu F_{\rho \sigma} = 0
    \label{eqnMR2}
\end{equation}
The $0$th component is:
\begin{align*}
    \epsilon^{0ijk} \partial_i F_{jk} &= -\epsilon_{ijk} \partial_i F_{jk} = 0 \\
    \implies \partial_iB_i &= 0
\end{align*}
and similarly the $i$th component gives \eqnref{eqnMEC}.

In summary, we find the covariant forms for Maxwell's Equations:
\begin{align*}
    \partial_\nu F^{\mu \nu} &= \mu_0 J^\mu \tag{\ref{eqnMR1}} \\
    \epsilon^{\mu \nu \rho \sigma} \partial_\nu F_{\rho \sigma} &= \partial_\nu (\hphantom{}^*F^{\mu \nu}) = 0 \tag{\ref{eqnMR2}}
\end{align*}
where we can introduce the notation of the tensor dual as seen in Ex3Q5.

The equation of motion for a charged particle is:
\begin{equation*}
    m\frac{dU^\mu}{d\tau} = f^{\mu} = q\eta^{\mu \nu} F_{\nu \sigma} U^\sigma
\end{equation*}
and charge conservation looks like:
\begin{equation}
    \partial_\mu J^\mu = 0
    \label{eqnRChargeConserve}
\end{equation}
If we write $F_{\mu \nu} = \partial_\mu A_\nu - \partial_\nu A_\mu$, we see that \eqnref{eqnMR2} is already satisfied, and \eqnref{eqnMR1} becomes:
\begin{equation*}
    \partial^\mu \partial_\nu A^\nu - \partial_\nu\partial^\nu A^\mu = \mu_0 J^\mu
\end{equation*}
which, in Lorentz gauge, becomes:
\begin{equation}
    -\Box A^\nu = \mu_0 J^\mu
    \label{eqnRPotPoisson}
\end{equation}
which gives the wave equation if $J = 0$, and is otherwise forced by charges and currents.
\end{document} 